---
title: "Correlation Matrix"
author: "Jonathan Huang"
date: "1/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Correlation Matrix

This document is geared toward making a correlation matrix between cells.
Cells are filtered based on the cells that are shown in the TSNE from the Seurat objects.
  
The particular variable "CellLabs", which is used to annotate the graphs, is pulled directly from the TSNE. The clusters listed in this file are determined by PCA.

Additional filtering is done to take certain genes out to stay parallel to the TSNE


```{r corsetup, echo=FALSE}
library(Hmisc)
library(data.table)
library(corrplot)
library(ComplexHeatmap)
library(dplyr)
library(circlize)
library(RColorBrewer)
library(googleVis)
## Color Palette
Tissue.Pal = c("lowgrade" = "#00BA38", "highgrade" = "#619CFF", "pdac"="#F8766D")
#sample.Pal = c("AM31" = "#B2DF8A", "AM86" = "#33A02C","AM48" = "#A6CEE3","AM72" = "#1F78B4","MK300"= "#FB9A99", "RS01" = "#E31A1C")
sample.Pal = c("lowgrade_IPMN_1" = "#B2DF8A", "lowgrade_IPMN_2" = "#33A02C","highgrade_IPMN_1" = "#A6CEE3","highgrade_IPMN_2" = "#1F78B4","pdac_1"= "#FB9A99", "pdac_2" = "#E31A1C")
cluster.Pal = c("0" = "#E5F5E0","1" ="royalblue", "2"="#FB6A4A", "3" ="#A1D99B" ,"4"="#FBB4AE","5"="#B3CDE3","6"="#CB181D","7"="#CCEBC5","8"="#31A354","9"="#DECBE4")

## Imports the TSNE files for the lists of cells that were not thrown out.
CellLabs <- read.csv("tsneLabelsv2.csv")
rownames(CellLabs) <- CellLabs$CellId
renameSamples <- function(x){
  if(x == "AM31"){
    return("lowgrade_IPMN_1")
  }
  if(x == "AM86"){
    return("lowgrade_IPMN_2")
  }
  if(x == "AM48"){
    return("highgrade_IPMN_1")
  }
  if(x == "AM72"){
    return("highgrade_IPMN_2")
  }
  if(x == "MK300"){
    return("pdac_1")
  }
  if(x == "RS01"){
    return("pdac_2")
  }
}
CellLabs$Sample <- sapply(CellLabs$Sample, renameSamples)



## The files to be added, names are self-explanatory

AM31 = as.data.table(read.csv('AM31-Combine-1to8_S1.counts.umiCounts.table.csv'))
AM31[, CellId := paste("AM31",CellId, sep="_")]
AM31 <- AM31[is.element(CellId, CellLabs$CellId)]
AM31dtmelt <- melt(AM31, id.vars="CellId")
# AM31dt[, value := ifelse(value > 0, 1, 0)]
AM31dt <- dcast(AM31dtmelt, variable ~ CellId)
AM31dt.null <- AM31dt[, variable:=NULL]
AM31dt.null[, rSums:=rowSums(AM31dt.null)]
AM31fdt <- AM31dt.null[rSums > 0]

AM48 = as.data.table(read.csv('AM48_Combine_S1.counts.umiCounts.aboveBackground.table.csv'))
AM48[, CellId := paste("AM48",CellId, sep="_")]
AM48 <- AM48[is.element(CellId, CellLabs$CellId)]
AM48dtmelt <- melt(AM48, id.vars="CellId")
# AM48dt[, value := ifelse(value > 0, 1, 0)]
AM48dt <- dcast(AM48dtmelt, variable ~ CellId)
AM48dt.null <- AM48dt[, variable:=NULL]
AM48dt.null[, rSums:=rowSums(AM48dt.null)]
AM48fdt <- AM48dt.null[rSums > 0]

MK300 = as.data.table(read.csv('MK300_Combine_S1.counts.umiCounts.aboveBackground.table.csv'))
MK300[, CellId := paste("MK300",CellId, sep="_")]
MK300 <- MK300[is.element(CellId, CellLabs$CellId)]
MK300dtmelt <- melt(MK300, id.vars="CellId")
# MK300dt[, value := ifelse(value > 0, 1, 0)]
MK300dt <- dcast(MK300dtmelt, variable ~ CellId)
MK300dt.null <- MK300dt[, variable:=NULL]
MK300dt.null[, rSums:=rowSums(MK300dt.null)]
MK300fdt <- MK300dt.null[rSums > 0]

RS01 = as.data.table(read.csv('RS01-Combined_S1.counts.umiCounts.aboveBackground.table.csv'))
RS01[, CellId := paste("RS01",CellId, sep="_")]
RS01 <- RS01[is.element(CellId, CellLabs$CellId)]
RS01dtmelt <- melt(RS01, id.vars="CellId")
# RS01dt[, value := ifelse(value > 0, 1, 0)]
RS01dt <- dcast(RS01dtmelt, variable ~ CellId)
RS01dt.null <- RS01dt[, variable:=NULL]
RS01dt.null[, rSums:=rowSums(RS01dt.null)]
RS01fdt <- RS01dt.null[rSums > 0]

AM72 = as.data.table(read.csv('AM72_Combine_S1.counts.umiCounts.aboveBackground.table.csv'))
AM72[, CellId := paste("AM72",CellId, sep="_")]
AM72 <- AM72[is.element(CellId, CellLabs$CellId)]
AM72dtmelt <- melt(AM72, id.vars="CellId")
AM72dt <- dcast(AM72dtmelt, variable ~ CellId)
AM72dt.null <- AM72dt[, variable:=NULL]
AM72dt.null[, rSums:=rowSums(AM72dt.null)]
AM72fdt <- AM72dt.null[rSums > 0]

AM86 = as.data.table(read.csv('AM86_combine_S1.counts.umiCounts.aboveBackground.table.csv'))
AM86[, CellId := paste("AM86",CellId, sep="_")]
AM86 <- AM86[is.element(CellId, CellLabs$CellId)]
AM86dtmelt <- melt(AM86, id.vars="CellId")
AM86dt <- dcast(AM86dtmelt, variable ~ CellId)
AM86dt.null <- AM86dt[, variable:=NULL]
AM86dt.null[, rSums:=rowSums(AM86dt.null)]
AM86fdt <- AM86dt.null[rSums > 0]


```


```{r process, echo=FALSE}
## creates correlation, joins to tsne clusters
corAM31 <- cor(AM31fdt[,rSums:=NULL], method="pearson")
#rcorAM31 <- rcorr(as.matrix(AM31fdt), type="pearson")
corAM31merge <- merge(corAM31, CellLabs, by=0, all.x=TRUE)
corAM31merge <- as.data.table(corAM31merge)
corAM31merge[,Row.names:=NULL]

corAM48 <- cor(AM48fdt[,rSums:=NULL], method="pearson")
#rcorAM48 <- rcorr(as.matrix(AM48fdt), type="pearson")
corAM48merge <- merge(corAM48, CellLabs, by=0, all.x=TRUE)
corAM48merge <- as.data.table(corAM48merge)
corAM48merge[,Row.names:=NULL]

corMK300 <- cor(MK300fdt[,rSums:=NULL], method="pearson")
#rcorMK300 <- rcorr(as.matrix(MK300fdt), type="pearson")
corMK300merge <- merge(corMK300, CellLabs, by=0, all.x=TRUE)
corMK300merge <- as.data.table(corMK300merge)
corMK300merge[,Row.names:=NULL]

corRS01 <- cor(RS01fdt[,rSums:=NULL], method="pearson")
#rcorRS01 <- rcorr(as.matrix(RS01fdt), type="pearson")
corRS01merge <- merge(corRS01, CellLabs, by=0, all.x=TRUE)
corRS01merge <- as.data.table(corRS01merge)
corRS01merge[,Row.names:=NULL]

corAM72 <- cor(AM72fdt[,rSums:=NULL], method="pearson")
corAM72merge <- merge(corAM72, CellLabs, by=0, all.x=TRUE)
corAM72merge <- as.data.table(corAM72merge)
corAM72merge[,Row.names:=NULL]

corAM86 <- cor(AM86fdt[,rSums:=NULL], method="pearson")
corAM86merge <- merge(corAM86, CellLabs, by=0, all.x=TRUE)
corAM86merge <- as.data.table(corAM86merge)
corAM86merge[,Row.names:=NULL]
```

## Plots
  
  
  
### Individiual Sample Plots

These plots consist of plots of each sample.


```{r plots, echo=FALSE}
# corrplot(corAM31, method="color", order="hclust")
exc <- c("Tissue.Type", "Cluster", "CellId", "Sample")
haAM31 <- HeatmapAnnotation(corAM31merge[,.(Tissue.Type, Sample, Cluster)], col = list(Tissue.Type = Tissue.Pal, Sample = sample.Pal, Cluster = cluster.Pal))
Heatmap(corAM31merge[,setdiff(names(corAM31merge), exc),with=FALSE], 
        name="correlation", 
        col=colorRamp2(c(0,1),c("white","blue")),
        show_row_names = FALSE, 
        show_column_names = FALSE,
        top_annotation = haAM31)

haAM48 <- HeatmapAnnotation(corAM48merge[,.(Tissue.Type, Sample, Cluster)],col = list(Tissue.Type = Tissue.Pal, Sample = sample.Pal, Cluster = cluster.Pal))
# corrplot(corAM48, method="color", order="hclust")
Heatmap(corAM48merge[,setdiff(names(corAM48merge), exc),with=FALSE], 
        name="correlation",
        col=colorRamp2(c(0,1),c("white","blue")), 
        show_row_names = FALSE,
        show_column_names = FALSE,
        top_annotation = haAM48)

haMK300 <- HeatmapAnnotation(corMK300merge[,.(Tissue.Type, Sample, Cluster)],col = list(Tissue.Type = Tissue.Pal, Sample = sample.Pal, Cluster = cluster.Pal))
# corrplot(corMK300, method="color", order="hclust")
Heatmap( corMK300merge[,setdiff(names(corMK300merge), exc),with=FALSE],
         name="correlation",
         col=colorRamp2(c(0,1),c("white","blue")),
         show_row_names = FALSE,
         show_column_names = FALSE,
         top_annotation = haMK300 )

haRS01 <- HeatmapAnnotation(corRS01merge[,.(Tissue.Type, Sample, Cluster)],col = list(Tissue.Type = Tissue.Pal, Sample = sample.Pal, Cluster = cluster.Pal))
# corrplot(corRS01, method="color", order="hclust")
Heatmap(corRS01merge[,setdiff(names(corRS01merge), exc),with=FALSE], 
        name="correlation", 
        col=colorRamp2(c(0,1),c("white","blue")),
        show_row_names = FALSE, 
        show_column_names = FALSE,
        top_annotation = haRS01)

haAM72 <- HeatmapAnnotation(corAM72merge[,.(Tissue.Type, Sample, Cluster)],col = list(Tissue.Type = Tissue.Pal, Sample = sample.Pal, Cluster = cluster.Pal))
Heatmap(corAM72merge[,setdiff(names(corAM72merge), exc),with=FALSE], 
        name="correlation", 
        col=colorRamp2(c(0,1),c("white","blue")),
        show_row_names = FALSE, 
        show_column_names = FALSE,
        top_annotation = haAM72)

haAM86 <- HeatmapAnnotation(corAM86merge[,.(Tissue.Type, Sample, Cluster)],col = list(Tissue.Type = Tissue.Pal, Sample = sample.Pal, Cluster = cluster.Pal))
Heatmap(corAM86merge[,setdiff(names(corAM86merge), exc),with=FALSE], 
        name="correlation", 
        col=colorRamp2(c(0,1),c("white","blue")),
        show_row_names = FALSE, 
        show_column_names = FALSE,
        top_annotation = haAM86)
```
  
  
  
### All Samples
  
  
Takes a look at all samples at once.
  
  
  
```{r bind, echo=FALSE}
## allsamp <- bind_rows(list(AM31, AM48, MK300, RS01))
## alldtmelt <- melt(allsamp, id.vars="CellId")
alldtmelt <- bind_rows(list(AM31dtmelt, AM48dtmelt, MK300dtmelt, RS01dtmelt, AM72dtmelt, AM86dtmelt))
alldtcast <- dcast(alldtmelt, variable ~ CellId, value.var = "value")
alldt.null <- alldtcast[, variable:=NULL]
```

```{r bindplot, echo=FALSE}
corall <- cor(alldt.null)
od = hclust(dist(corall))$order
corall = corall[od, od]
corallmerge <- merge(corall, CellLabs, by=0, all.x=TRUE)
corallmerge <- as.data.table(corallmerge)
corallmerge[, Cluster := as.factor(Cluster)]
corallmerge[,Row.names:=NULL]

haall <- HeatmapAnnotation(corallmerge[,.(Tissue.Type, Sample, Cluster)], col = list(Tissue.Type = Tissue.Pal, Sample = sample.Pal, Cluster = cluster.Pal))
                           # annotation_legend_param = list(V3 = list(
                           #   title = "Cluster",
                           #   labels = c(0,1,2,3,4,5,6,7,8,9)
                           # )))
# corrplot(corRS01, method="color", order="hclust")
Heatmap(corallmerge[,setdiff(names(corallmerge), exc),with=FALSE], 
        name="correlation", 
        col=colorRamp2(c(0,1),c("white","blue")),
        show_row_names = FALSE, 
        show_column_names = FALSE,
        top_annotation = haall)
```

### Cluster by Cluster

Look at individual clusters (e.g. cluster 0, cluster 3 etc).

```{r clusters, echo=FALSE}

## have to recreate the correlations
lapply(c(0:9), function(x){
  clustercells <- subset(CellLabs, Cluster == x)
  dt <- alldtmelt[is.element(CellId, clustercells$CellId)]
  dt <- dcast(dt, variable ~ CellId, value.var = "value")
  corclust <- cor(dt[, variable:=NULL])
  od = hclust(dist(corclust))$order
  corclust = corclust[od, od]
  corclustmerge <- merge(corclust, clustercells, by=0, all.x=TRUE)
  corclustmerge <- as.data.table(corclustmerge)
  corclustmerge[,Row.names:=NULL]
  
  ## plot
  haall <- HeatmapAnnotation(corclustmerge[,.(Tissue.Type, Sample)], col = list(Tissue.Type = Tissue.Pal, Sample = sample.Pal))
  Heatmap(corclustmerge[,setdiff(names(corclustmerge), exc),with=FALSE], 
        name="correlation", 
        col=colorRamp2(c(0,1),c("white","blue")),
        show_row_names = FALSE, 
        show_column_names = FALSE,
        top_annotation = haall)
  # print(paste("Cluster: ", x))
})

```

### Manual Clusters

```{r clustman, echo=FALSE}
clustercells <- subset(CellLabs, Cluster == 9 | Cluster == 7)
dt <- alldtmelt[is.element(CellId, clustercells$CellId)]
dt <- dcast(dt, variable ~ CellId, value.var = "value")
corclust <- cor(dt[, variable:=NULL])
od = hclust(dist(corclust))$order
corclust = corclust[od, od]
corclustmerge <- merge(corclust, clustercells, by=0, all.x=TRUE)
corclustmerge <- as.data.table(corclustmerge)
corclustmerge[, Cluster := as.factor(Cluster) ]
corclustmerge[,Row.names:=NULL]
  
## plot
haall <- HeatmapAnnotation(corclustmerge[,.(Tissue.Type, Sample)], col = list(Tissue.Type=Tissue.Pal, Sample = sample.Pal))
Heatmap(corclustmerge[,setdiff(names(corclustmerge), exc),with=FALSE], 
        name="correlation", 
        col=colorRamp2(c(0,1),c("white","blue")),
        show_row_names = FALSE, 
        show_column_names = FALSE,
        top_annotation = haall)
```

### UMI and Gene count by Cluster
  
  
UMI counts as a total per sample and genes per sample as well.
For UMI, since each nUMI for each gene is far too many points, instead, this will be a sum and mean of counts per cell. 

e.g. cell x in cluster y has a mean of n umi and has a total of m umi etc.

note: median isn't used since it's always 0. Mean is number of Sum/nGene

```{r counts, echo=FALSE, include=FALSE}
## Add counts and median and mean per cell
# AM31
AM31[, c("nsum", "nGene") := .(rowSums(AM31[,2:ncol(AM31)]),
                               Reduce('+', lapply(AM31[,2:ncol(AM31)], function(x) ifelse(x > 0, 1, 0)))
                              )]
AM31[, nmean := nsum/nGene]
AM31counts <- AM31[, .(CellId, nsum, nmean, nGene)]

# AM48
AM48[, c("nsum", "nGene") := .(rowSums(AM48[,2:ncol(AM48)]),
                               Reduce('+', lapply(AM48[,2:ncol(AM48)], function(x) ifelse(x > 0, 1, 0)))
                              )]
AM48[, nmean := nsum/nGene]
AM48counts <- AM48[, .(CellId, nsum, nmean, nGene)]

# MK300
MK300[, c("nsum", "nGene") := .(rowSums(MK300[,2:ncol(MK300)]),
                               Reduce('+', lapply(MK300[,2:ncol(MK300)], function(x) ifelse(x > 0, 1, 0)))
                              )]
MK300[, nmean := nsum/nGene]
MK300counts <- MK300[, .(CellId, nsum, nmean, nGene)]

# RS01
RS01[, c("nsum", "nGene") := .(rowSums(RS01[,2:ncol(RS01)]),
                               Reduce('+', lapply(RS01[,2:ncol(RS01)], function(x) ifelse(x > 0, 1, 0)))
                              )]
RS01[, nmean := nsum/nGene]
RS01counts <- RS01[, .(CellId, nsum, nmean, nGene)]

# AM72
AM72[, c("nsum", "nGene") := .(rowSums(AM72[,2:ncol(AM72)]),
                               Reduce('+', lapply(AM72[,2:ncol(AM72)], function(x) ifelse(x > 0, 1, 0)))
                              )]
AM72[, nmean := nsum/nGene]
AM72counts <- AM72[, .(CellId, nsum, nmean, nGene)]

# AM86
AM86[, c("nsum", "nGene") := .(rowSums(AM86[,2:ncol(AM86)]),
                               Reduce('+', lapply(AM86[,2:ncol(AM86)], function(x) ifelse(x > 0, 1, 0)))
                              )]
AM86[, nmean := nsum/nGene]
AM86counts <- AM86[, .(CellId, nsum, nmean, nGene)]

allsum <-  bind_rows(list(AM31counts, AM48counts, MK300counts, RS01counts, AM72counts, AM86counts))
```

```{r clusternplotplots, echo=FALSE, include=FALSE}
allsum <- merge(allsum, CellLabs, on="CellId")
```