---
title: "CTC Analysis"
author: "Jonathan Huang"
date: "12/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of CTC

This document covers the analysis of our CTC's.

## Setup

CTCS are organized in a table format where one column corresponds to a single biomarker in that cell.
E.g.

```{r ctcsetup, echo=FALSE}
library(stats)
library(data.table)
library(plyr)
library(dplyr)
library(survival)
library(survminer)
library(ggplot2)
library(ggpubr)
library(stargazer)
CTCdt <- fread('CTCcelllistFull.csv')
drawlist <- CTCdt[,.N,by=Draw.ID]
stargazer(head(CTCdt), type="text", summary=rep(FALSE, length(head(CTCdt))))
CTCdemo <- fread('CTCdemo.csv')
setkey(CTCdemo, Patient.ID)
CTCdemo <- unique(CTCdemo)
CTCdemo[,c('T','N','M','Stage') := .(gsub('[cp]','',T),
                                     gsub('[cp]','',N),
                                     gsub('[cp]','',M),
                                     gsub('[AB]','',Stage)
                                     )
        ]
CTCCA19 <- fread('CTCCA19.csv')
CTCCA19[,CA19 := as.numeric(CA19)]
setkey(CTCCA19, Patient.ID)
CTCCA19 <- unique(CTCCA19)
CTCCA19[, CA19_date := as.Date(draw_date)]
sampledates <- fread('CTCsamples.csv')
setkey(sampledates, Draw.ID)
sampledates <- unique(sampledates)
imagedates <- fread('ctcrecist.csv')
imagedates[, image_date := as.Date(draw_date)]
therapy <- fread('CTCTherapies.csv')
therapy[, c('start', 'end') := .(as.Date(start), as.Date(end))]
setkeyv(therapy, c("Patient.ID","start"))
tstart <- therapy[, .SD[1], by=.(Patient.ID,line)]
tstart <- tstart[, .(Patient.ID,line,start)]
tstartdates <- tstart[!is.na(start), .SD[1], by=Patient.ID]
setnames(tstartdates, old='start', new='t.start')
tstartdates <- tstartdates[, .(Patient.ID,t.start)]
tend <- therapy[, .SD[.N], by=.(Patient.ID,line)]
tend <- tend[, .(Patient.ID,line,end)]
therapy <- tstart[tend, on=.(Patient.ID,line)]
therapy[, draw_date := as.Date(start)]
```
These will be organized based on Cell.ID with Biomarkers as a list.

```{r ctcsetup2, echo=FALSE, warning=FALSE}
EM = function(x){
  output = ''
  if(is.element(1, x) || is.element(4, x)){
    if(is.element(3, x)){
      return('SC')
    } else {
      output = paste(output,'E', sep="")
    }

  }
  if(is.element(2, x)){
    output = paste(output,'M', sep="")
  }
  return(output)
}
Epc = function(x){
  output=''
  if(is.element(2, x)){
    output= paste(output,'m',sep="")
  }
  if(is.element(1, x)){
    output = paste(output,'p',sep="")
  }
  if(is.element(4, x)){
    output = paste(output,'c',sep='')
  }
  return(output)
}
CTCdt2 <- CTCdt[,.(isCTC, Biomarkers = list(Biomarker)), by=.(Cell.ID, isCTC,Draw.ID, Patient.ID)]
DT <- CTCdt2[isCTC==1,.(Patient.ID, Draw.ID, Cell.ID, isCTC, Biomarkers, cellType = lapply(Biomarkers, function(x){EM(x)}), epType = lapply(Biomarkers, function(x){Epc(x)}))]
DT[, c('cellType', 'epType') := .(paste(cellType, sep=""),paste(epType, sep=""))]
mDT <- melt(DT[,.N,by=.(Patient.ID,Draw.ID,isCTC,cellType)], id=c('Patient.ID', 'Draw.ID','isCTC','cellType'))
temp <- mDT
mDT <- dcast(mDT, Patient.ID + Draw.ID + isCTC ~ cellType)
temp2 <- mDT
## add draws with no CTCs as 0's
adding <- merge(drawlist, mDT, by='Draw.ID', all=TRUE)
adding[is.na(Patient.ID), Patient.ID := gsub('-.*','',Draw.ID)]
adding[is.na(isCTC), isCTC:=1]
mDT <- adding[(Draw.ID != '' & !grepl('.*-[T].*',Draw.ID))]
mDT <- sampledates[mDT, on='Draw.ID']
mDT2 <- melt(DT[,.N,by=.(Patient.ID,Draw.ID,isCTC,epType)], id=c('Patient.ID', 'Draw.ID','isCTC','epType'))
mDT2 <- dcast(mDT2, Patient.ID + Draw.ID + isCTC ~ epType)
mDT2 <- mDT2[,.(Draw.ID,isCTC,p,c,pc,m,mp,mc,mpc)]
DT2 <- mDT[CTCdemo, on='Patient.ID']
DT2 <- mDT2[DT2, on=c('Draw.ID','isCTC')]
DT3 <- CTCCA19[DT2, on=.(Patient.ID,draw_date), roll=-Inf]
DT3[,c('E','EM','M','SC','p','c','pc','m','mp','mc','mpc')][is.na(DT3[,c('E','EM','M','SC','p','c','pc','m','mp','mc','mpc')])] <- 0
DT4 <- imagedates[DT3, on=.(Patient.ID,draw_date), roll=-Inf]
DT4[,draw_date := as.Date(draw_date)]
setkey(DT4, Draw.ID)
DT4 <- unique(DT4)
# removed M as we cant be positive that they are CTCs
DT4[, totCTC := sum(E + EM + SC), by=Draw.ID]

# set up proportions
DT4[, c('perE', 'perEM', 'perSC') := .(E/totCTC*100, EM/totCTC*100, SC/totCTC*100), by=Draw.ID]

# therapies added
DTT <- tstartdates[DT4, on=.(Patient.ID)]
DTT <- therapy[DTT, on=.(Patient.ID, draw_date), roll=TRUE]
DTT[, c("days.first", "days.start") := .(draw_date - t.start, draw_date - start)]
```


## Demographics

```{r demo, echo=FALSE}
colheads <- c('Gender','Stage','T','N','M')
demooccur <- lapply(colheads, function(x){
  CTCdemo[,.N,by=x]
})
stargazer(demooccur, type='text',summary=rep(FALSE, length(demooccur)))
```

## Compare CTC counts in Baseline Treatment Naive patient groups

```{r counts, echo=FALSE, warning=FALSE}
BTNDT <- DT4[grepl('.*-1$', Draw.ID) & BTN == 1]
stargazer(BTNDT[isCTC==1,.(E,EM,M)], type='text', title='Distribution of CTC cells')
## stargazer(BTNDT[isCTC==0,.(E,EM,M)], type='text', title='Distribution of Normal cells')
p <- ggplot(BTNDT[isCTC==1]) + geom_boxplot(aes(x='Epithelial', y=E), group=1) + geom_jitter(aes(x='Epithelial', y=E), group=1) + 
                     geom_boxplot(aes(x='Transition', y=EM), group=2) + geom_jitter(aes(x='Transition', y=EM), group=2) +
                     geom_boxplot(aes(x='Mesenchymal', y=M), group=3) + geom_jitter(aes(x='Mesenchymal', y=M),group=3) +
    labs(title="General Distribution between Epithelial and Mesenchymal Cells in Baseline Treatment Naive Samples", y='Number of Cells', x='')

plot(p)


```

### Counts

```{r counts2, echo=FALSE, warning=FALSE}
colheads <- c('Gender','Stage','T','i.N','i.M')
btnoccur <- lapply(colheads, function(x){
  BTNDT[,.(.N, sumE = sum(E), sumEM = sum(EM), sumM = sum(M), total = sum(E+M+EM)), by=x]
})
stargazer(btnoccur, type='text',summary=rep(FALSE, length(btnoccur)))
```

### Distributions
```{r distr, echo=FALSE, warning=FALSE}
colheads <- c('Gender','Stage','i.M')
j <- c('E','EM','M','SC','Gender','Stage','Tumor.Type','i.M')
plotdistri <- function(x,y,m){
  temp <- melt(x[,m, with=FALSE])
  sapply(y, function(n){
          p <-ggplot(temp, aes(x=variable)) +
            geom_jitter(aes_string(y='value',color=n),position = position_jitter(width = .4), group=1) + geom_boxplot(aes_string(y='value',color=n),alpha=.01, group=1) +
            geom_jitter(aes_string(y='value',color=n),position = position_jitter(width = .4), group=2) + geom_boxplot(aes_string(y='value',color=n),alpha=.01, group=2) +
            geom_jitter(aes_string(y='value',color=n),position = position_jitter(width = .4),group=3) + geom_boxplot(aes_string(y='value',color=n),alpha=.01, group=3) +
            geom_jitter(aes_string(y='value',color=n),position = position_jitter(width = .4),group=4) + geom_boxplot(aes_string(y='value',color=n),alpha=.01, group=4)
      plot(p)
  })
}
plotdistri(BTNDT, colheads,j)
```

#### EMT breakdowns

```{r distrb, echo=FALSE, message=FALSE, warning=FALSE}
jb <- c('p','c','pc','mp','mc','mpc','Gender','Stage','Tumor.Type','T','i.N','i.M')
plotdistri(BTNDT,colheads,jb)
```
### Percent Distributions

```{r percentdistr, echo=FALSE, message=FALSE, warning=FALSE}
colheads <- c('Gender','Stage','i.M')
j <- c('E','EM','M','SC','Gender','Stage','Tumor.Type','i.M')
emtper <- BTNDT[, .(Patient.ID,E = E /(E + EM + M + SC), EM = EM/(E+EM+M + SC), M = M/(E+EM+M+SC), SC=SC/(E+EM+M+SC), Gender, Stage,i.M, Tumor.Type,vital.status,OS,Progression,PFS)]
plotdistri(emtper, colheads, j)

```

## Longitudinal Observations

```{r longsetup, echo=FALSE, message=FALSE, warning=TRUE}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

longFrames <- function(x){
  ## x is datatable
  setkey(x, Patient.ID)
  patlist <- unique(x[,Patient.ID])
  ## patgroups <- lapply(patlist, function(y){x[x$Patient.ID == y,]})
  patgroups <- lapply(patlist, function(y){ x[Patient.ID == y] })
  return(patgroups)
}

plotLong <- function(n){
  ## x is datatable/list of datatable
  cols <- c("Epithelial"='#e41a1c', "EMT"='#377eb8',"Mesenchymal"='#4daf4a', "CA19"='#984ea3', "Stem Cells" = "#ff7f00")
  ps <- lapply(n, function(x){
    ggplot(x, aes(x=Draw.ID, y=E,  color="Epithelial")) + 
      geom_point(group=1) + geom_line(group=1) + 
      geom_point(aes(x=Draw.ID, y=EM, color="EMT"), group=2) + geom_line(aes(x=Draw.ID, y=EM, color="EMT"),group=2) +
      geom_point(aes(x=Draw.ID, y=M, color="Mesenchymal"), group=3) + geom_line(aes(x=Draw.ID, y=M, color="Mesenchymal"),group=3) + 
      geom_point(aes(x=Draw.ID, y=SC, color="Stem Cells"), group=5) + geom_line(aes(x=Draw.ID, y=SC, color="Stem Cells"),group=5) +
      geom_point(aes(x=Draw.ID, y=-1, color=response), size=3, group=4)+
      labs(x="Draw ID", y="") + theme(legend.position = 'bottom')
  })
  ca19 <- lapply(n, function(x){
    ggplot(x, aes(x=Draw.ID, y=CA19, color="CA19")) +
      geom_point(group=1) + geom_line(group=1) +
      geom_point(aes(x=Draw.ID, y=-1, color=response), size=3, group=4)+
      labs(x="Draw ID", y="CA19-9") + theme(legend.position = 'bottom')
  })
  mapply(function(x,y){multiplot(x,y,cols=2)},ps,ca19)
}

```

```{r temp, echo=FALSE}
firstPD <- function(x){
  # takes in data.table
  temp <- x[response == "PD", .SD[1], by=Patient.ID]
  temp[, pred := 1]
  temp <- temp[, .(Draw.ID,pred)]
  joined <- temp[x, on=Draw.ID]
  return(joined)
}

lastNotPD <- function(x){
  # takes in DTT which then applys longFrames to. 
  temp <- longFrames(x)
  PDlist <- lapply(temp, function(x){
    t1 <- as.data.table(x)
    if((!"PD" %in% x$response)){
      x[, pred := 1]
      return(x[response != "PD" & !is.na(response), .SD[.N]])
    }
  })
  return(PDlist)
}

```

```{r longplot, echo=FALSE, message=FALSE, warning=FALSE}
groups <- longFrames(DT4)
plotLong(groups)
```

```{r predictionsetup, echo=FALSE, message=FALSE}
DT5 <- DT4
setkey(DT5, Draw.ID)
DT5 <- unique(DT5)
DT5 <- DT5[order(Patient.ID, draw_date)]
addSinglePred <- function(x){
  # function takes in a list of datatables and applies adds a day of (response) and next response (next) as columns. If response doesn't exist put in NA
  tDT <- x[, c('rnext', 'rdate', 'rnext2', 'rdate2') := .(shift(response, n=1L, fill=NA, type="lead"),
                                            shift(draw_date, type="lead") - draw_date,
                                            shift(response, n=2L, fill=NA, type="lead"),
                                            shift(draw_date, n=2L, type="lead") - draw_date), by=Patient.ID]
  return(tDT)
}
addTrend <- function(x){
  tDT <- x[, c('tE', 'tEM', 'tM', 'tSC', 'tnext','tdate','tnext2','tdate2') := .(
                                              E - shift(E),
                                              EM - shift(EM),
                                              M - shift(M),
                                              SC - shift(SC),
                                              shift(response, n=1L, fill=NA, type="lead"),
                                              shift(draw_date, type="lead") - draw_date,
                                              shift(response, n=2L, fill=NA, type="lead"),
                                              shift(draw_date, n=2L, type="lead") - draw_date
    ), by=Patient.ID]
  tDT <- x[, c('tpE', 'tpEM', 'tpSC') := .(
                                              perE - shift(perE),
                                              perEM - shift(perEM),
                                              perSC - shift(perSC)
    ), by=Patient.ID]
  return(tDT)
}

addSinglePred(DT5)
addTrend(DT5)

processPredictions <- function(x){
  tDT <- x[, rProg := mapply(function(a, b, c, d, e){
        if(!is.na(a)){
          if(a == 'PD'){
            return(1L)
          }
        }
        if (!is.na(d)) {
          if(b == 'PD'  && d <= 180){
            return(1L)
          }
        }
        if (!is.na(e)) {
          if(c == 'PD' && e <= 180){
            return(1L)            
          }
        }
        return(0L)
      
  },response, rnext, rnext2, rdate, rdate2)
    ]
  
  tDT <- tDT[, tProg := mapply(function(a, b, c, d, e){
        if(!is.na(a)){
          if(a == 'PD'){
            return(1L)
          }
        }
        if (!is.na(d)) {
          if(b == 'PD'  && d <= 180){
            return(1L)
          }
        }
        if (!is.na(e)) {
          if(c == 'PD' && e <= 180){
            return(1L)            
          }
        }
        return(0L)
      
  },response, tnext, tnext2, tdate, tdate2)
    ]
  return(tDT)
}

```
```{r test}

imageAsPoint <- function(x){
  tDT <- x
  tDT[, Draw.Index := gsub('.*-', '', Draw.ID)]
  tDT[,sList := mapply(function(x,y,z){
                            l <- list( )
                            return(l)
    }, SC, draw_date, Draw.Index)]
  return(tDT)
}
```
