---
title: "CTC ROC and AUC + more"
author: "Jonathan Huang"
date: "3/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ROCs and Thresholds

The purpose of this document is to attempt to uncover a strong relationship between variables and outcome where the variable is binary from a threshold.

To do so, there are several directions to take the data. The two general overarching methods are Sample based and Patient Based.

Sample Based
> Sample based is more concerned with samples in general. In this case, more samples are used. But the issue is that similar/different events can have both the same and different outcomes. This tends to lead to an issue where there is no statistical significance at all. Of course this issue is less apparent when the correlation is very strong. Tends to cover
    * One sample
    * Delta between samples
  Reliant on time as a distinct variable. E.g. 3 or 6 months from the point is the time limit.

Patient Based
> Patient based is more "longitudinal". Each Patient represents a point, variables are dependent a single event which can be
    * Means
    * Trends
    * One of Many
    etc.
  Primarily concerned with determining if a patient has an event or not.
    

```{r datainput, echo=FALSE, message=FALSE}
library(data.table)
library(plyr)
library(dplyr)
library(survival)
library(survminer)
library(ggplot2)
library(ggpubr)
library(stargazer)
library(stringr)
library(pROC)
library(ROCR)
library(scales)
ontherapy <- fread("ontherapy.csv")

####################
# ALL CTC
allct <- fread("AllCTCwnormals.csv")
# set variables as dates
allct[, c("draw_date", "t.start") := .(as.Date(draw_date), as.Date(t.start))]
allct <- ontherapy[allct, on="Draw.ID"]
####################
# TN only
TN <- fread("CTCdataNormalized.csv")
TN[, c("draw_date", "t.start") := .(as.Date(draw_date), as.Date(t.start))]
TN <- ontherapy[TN, on="Draw.ID"]
```


## Sample Based

Looking at samples.

```{r sample, echo=FALSE}
# Make copies to edit
sampA <- allct[i.M != "Mx"]
sampB <- TN 
```


### Predict 3 or 6 months

Look at samples as points. Given a value, does the patient have a PD within 3/6 months of that draw? Can be looked at with or without taking therapy into account.
note: uses Days instead of Months. e.g. 90 days vs 180 days. Ignores BTN draws
Issue may arise from the fact that the dates for response are the corresponding to draw_dates vs image dates.

```{r samp36, message=FALSE}
predictMonths <- function(dt, days, s, therapy){
  temp <- dt[Patient.ID == str_replace(eval(s), "-.*", ""),
             .(Patient.ID, Draw.ID, draw_date, t.start, response)]
  temp <- temp[draw_date <= (temp[Draw.ID == eval(s), draw_date] + days)]
  if(therapy == 1){
    temp <- temp[t.start == temp[Draw.ID == eval(s), t.start] & response == "PD",.N]
    ifelse(temp >= 1, return(1), return(0))
  } else {
    temp <- temp[response == "PD",.N]
    ifelse(temp >= 1, return(1), return(0))
  }
}
predictMonthsPR <- function(dt, days, s, therapy){
  temp <- dt[Patient.ID == str_replace(eval(s), "-.*", ""),
             .(Patient.ID, Draw.ID, draw_date, t.start, response)]
  temp <- temp[draw_date <= (temp[Draw.ID == eval(s), draw_date] + days)]
  if(therapy == 1){
    temp <- temp[t.start == temp[Draw.ID == eval(s), t.start] & response == "PR",.N]
    ifelse(temp >= 1, return(1), return(0))
  } else {
    temp <- temp[response == "PD",.N]
    ifelse(temp >= 1, return(1), return(0))
  }
}
deltaDates <- function(x){
  # function takes in a list of datatables and applies adds a day of (response) and next response (next) as columns. If response doesn't exist put in NA
  tDT <- x[, c('rnext', 'rdate') := .(shift(response, n=1L, fill=NA, type="lead"),
                                            shift(draw_date, type="lead") - draw_date)
                                            , by=Patient.ID]
  return(tDT)
}
simpDelta <- function(x, vars){
  tDT <- x
  for(i in 1:length(vars)){
    var = vars[i]
    tvar = paste("t",var,sep="")
    tDT[, eval(tvar) := get(var) - shift(get(var)), by=.(Patient.ID, line)]
  }
  return(tDT)
}
compDelta <- function(x, vars){
  tDT <- x
  for(i in 1:length(vars)){
    var = vars[i]
    tvar = paste("p",var,sep="")
    tDT[, eval(tvar) := get(var)/(shift(get(var)) + 1), by=.(Patient.ID, line)]
  }
  return(tDT)
}

plotdistri <- function(x,y,m, pred){
  temp <- melt(x[,m, with=FALSE], id.vars=c(y, pred))
  temp[, value:= ifelse(value > 200, 200,value)]
  sapply(y, function(n){
          # p.means <- ddply(temp, c(n, pred), summarise, mn = mean(value))
          p <- ggplot(temp, aes(x=factor(get(pred)))) +
            geom_point(aes_string(y='value',color=pred),position = position_jitterdodge(dodge.width = .7), group=1) + geom_boxplot(aes_string(y='value',fill=pred),alpha=.01, group=1) +
            stat_compare_means(aes_string(y='value', group=pred), label="p.format") +
            facet_grid(get(n)~variable, scales="free") +
            labs(title=paste("Distribution of Variables dependent on M staging: ",pred), x=pred)
            # geom_text(data = p.means, aes(x=factor(get(pred)), y=mn, label=mn), size=5, vjust=-1.5)
      plot(p)
  })
}
plotdistriNS <- function(x,y,m, pred){
  temp <- melt(x[,m, with=FALSE], id.vars=c(y, pred))
  temp[, value:= ifelse(value > 200, 200,value)]
  sapply(y, function(n){
          p <- ggplot(temp, aes(x=factor(get(pred)))) +
            geom_point(aes_string(y='value',color=pred),position = position_jitterdodge(dodge.width = .7), group=1) + geom_boxplot(aes_string(y='value',fill=pred),alpha=.01, group=1) +
            stat_compare_means(aes_string(y='value', group=pred), label="p.format") +
            facet_grid(.~variable, scales="free") +
            labs(title=paste("Distribution of Variables dependent on M staging: ",pred), x=pred)
      plot(p)
  })
}
####### 
# Allct
sampA[, c("pred3", "pred6") := .(sapply(Draw.ID, function(x){predictMonths(sampA, 90, x, 1)}),
                                 sapply(Draw.ID, function(x){predictMonths(sampA, 180, x, 1)}))]
sampA[, c("pred3", "pred6") := .(as.factor(pred3), as.factor(pred6))]
sampA <- deltaDates(sampA)
Aon <- simpDelta(sampA[On.Therapy != 0 & response != "Baseline untreated"], c("E", "EM", "M", "SC", "tot","cfKRAS","exoKRAS"))
Aon <- compDelta(Aon, c("E", "EM", "M", "SC", "tot"))
Aline <- simpDelta(sampA[response != "Baseline untreated"], c("E", "EM", "M", "SC", "tot", "ve","vc","vec","cfKRAS","exoKRAS"))
Aline <- compDelta(Aline, c("E", "EM", "M", "SC", "tot","ve","vc","vec"))

#######
# BTN
sampB[, c("pred3", "pred6", "pred4", "pred8") := .(sapply(Draw.ID, function(x){predictMonths(sampB, 90, x, 1)}),
                                 sapply(Draw.ID, function(x){predictMonths(sampB, 180, x, 1)}),
                                 sapply(Draw.ID, function(x){predictMonths(sampB, 120, x, 1)}),
                                 sapply(Draw.ID, function(x){predictMonths(sampB, 240, x, 1)}))]
sampB[, c("pred3", "pred6", "pred4", "pred8") := .(as.factor(pred3), as.factor(pred6), as.factor(pred4), as.factor(pred8))]

## samp months PR
sampB[, c("PRpred3", "PRpred6", "PRpred4", "PRpred8") := .(sapply(Draw.ID, function(x){predictMonthsPR(sampB, 90, x, 1)}),
                                 sapply(Draw.ID, function(x){predictMonthsPR(sampB, 180, x, 1)}),
                                 sapply(Draw.ID, function(x){predictMonthsPR(sampB, 120, x, 1)}),
                                 sapply(Draw.ID, function(x){predictMonthsPR(sampB, 240, x, 1)}))]
sampB[, c("PRpred3", "PRpred6", "PRpred4", "PRpred8") := .(as.factor(pred3), as.factor(pred6), as.factor(pred4), as.factor(pred8))]
Bon <- simpDelta(sampB[On.Therapy != 0 & response != "Baseline untreated"], c("prop.E", "prop.EM","prop.SC","prop.tot"))
Bon <- compDelta(Bon, c("E", "EM", "SC","tot"))

A <- sampA[response != "Baseline untreated"]
B <- sampB[response != "Baseline untreated"]
BTN <- sampB[response == "Baseline untreated"]
BTN[, Progression := as.factor(Progression)]

```



#### General Distributions based on Prediction in 3/6 months

Do we see a difference between the points that have shown progression in the past 3-6 months?

##### All Patients

```{r pred36distALL, echo=FALSE, include=FALSE}
## Allct
# j <- c('tot', 'totU','i.M',"pred3")
# plotdistri(A, c("i.M"), j, "pred3")
# j <- c('tot', 'totU','i.M',"pred6")
# plotdistri(A, c("i.M"), j, "pred6")
# 
# j <- c('E','EM','M','SC','numClust','i.M', "pred3")
# plotdistri(A,c("i.M"),j, "pred3")
# j <- c('E','EM','M','SC','numClust','i.M', "pred6")
# plotdistri(A, c("i.M"), j, "pred6")
# 
# j <- c('e','c','ec','ve','vc','vec','i.M',"pred3")
# plotdistri(A,c("i.M"),j, "pred3")
# j <- c('e','c','ec','ve','vc','vec','i.M',"pred6")
# plotdistri(A, c("i.M"), j, "pred6")
# 
# ## looking at groupings together
# j <- c('e2', 've2', 'vec', 've.vc', 'i.M', 'pred3')
# plotdistri(A,c("i.M"),j, "pred3")
# j <- c('e2', 've2', 'vec', 've.vc', 'i.M', 'pred6')
# plotdistri(A,c("i.M"),j, "pred6")

# j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'pred3')
# plotdistri(A,c("i.M"),j, "pred3")
# j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'pred6')
# plotdistri(A,c("i.M"),j, "pred6")
# 
# j <- c('CA19', 'i.M', 'pred3')
# plotdistri(A,c("i.M"),j, "pred3")
# j <- c('CA19', 'i.M', 'pred6')
# plotdistri(A,c("i.M"),j, "pred6")
# 
# ## draw to draw
# j <- c('tE', 'tEM', 'tM', 'tSC', 'ttot', 'i.M', 'pred3')
# plotdistri(Aon,c("i.M"),j, "pred3")
# j <- c('tE', 'tEM', 'tM', 'tSC', 'ttot', 'i.M', 'pred6')
# plotdistri(Aon,c("i.M"),j, "pred6")
# 
# j <- c('pE', 'pEM', 'pM', 'pSC', 'ptot', 'i.M', 'pred3')
# plotdistri(Aon,c("i.M"),j, "pred3")
# j <- c('pE', 'pEM', 'pM', 'pSC', 'ptot', 'i.M', 'pred6')
# plotdistri(Aon,c("i.M"),j, "pred6")

### same line doesnt matter if on therapy (no btn)
j <- c('tE', 'tEM', 'tM', 'tSC', 'ttot', 'i.M', 'pred3')
plotdistri(Aline,c("i.M"),j, "pred3")
j <- c('tE', 'tEM', 'tM', 'tSC', 'ttot', 'i.M', 'pred6')
plotdistri(Aline,c("i.M"),j, "pred6")

j <- c('pE', 'pEM', 'pM', 'pSC', 'ptot', 'i.M', 'pred3')
plotdistri(Aline,c("i.M"),j, "pred3")
j <- c('pE', 'pEM', 'pM', 'pSC', 'ptot', 'i.M', 'pred6')
plotdistri(Aline,c("i.M"),j, "pred6")

j <- c('tve', 'tvc', 'tvec', 'i.M', 'pred3')
plotdistri(Aline,c("i.M"),j, "pred3")
j <- c('tve', 'tvc', 'tvec', 'i.M', 'pred6')
plotdistri(Aline,c("i.M"),j, "pred6")

j <- c('pve', 'pvc', 'pvec', 'i.M', 'pred3')
plotdistri(Aline,c("i.M"),j, "pred3")
j <- c('pve', 'pvc', 'pvec', 'i.M', 'pred6')
plotdistri(Aline,c("i.M"),j, "pred6")
```


##### BTN

```{r justpoints, echo=FALSE, include=FALSE}
# j <- c('tot', 'totU','i.M',"pred3")
# plotdistri(B, c("i.M"), j, "pred3")
# j <- c('tot', 'totU','i.M',"pred6")
# plotdistri(B, c("i.M"), j, "pred6")
# 
# j <- c('E','EM','M','SC','numClust','i.M', "pred3")
# plotdistri(B,c("i.M"),j, "pred3")
# j <- c('E','EM','M','SC','numClust','i.M', "pred6")
# plotdistri(B, c("i.M"), j, "pred6")
# 
# 
# j <- c('e','c','ec','ve','vc','vec','i.M',"pred3")
# plotdistri(B,c("i.M"),j, "pred3")
# j <- c('e','c','ec','ve','vc','vec','i.M',"pred6")
# plotdistri(B, c("i.M"), j, "pred6")
# 
# ## looking at groupings together
# j <- c('e2', 've2', 'vec', 'i.M', 'pred3')
# plotdistri(B,c("i.M"),j, "pred3")
# j <- c('e2', 've2', 'vec', 'i.M', 'pred6')
# plotdistri(B,c("i.M"),j, "pred6")

j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'pred3')
plotdistri(B,c("i.M"),j, "pred3")
j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'pred6')
plotdistri(B,c("i.M"),j, "pred6")

j <- c('CA19', 'i.M', 'pred3')
plotdistri(B,c("i.M"),j, "pred3")
j <- c('CA19', 'i.M', 'pred6')
plotdistri(B,c("i.M"),j, "pred6")

## draw to draw
j <- c('pE', 'pEM', 'pSC', 'ptot', 'i.M', 'pred3')
plotdistri(Bon,c("i.M"),j, "pred3")
j <- c('pE', 'pEM', 'pSC', 'ptot', 'i.M', 'pred6')
plotdistri(Bon,c("i.M"),j, "pred6")

j <- c('tprop.E', 'tprop.EM', 'tprop.SC', 'tprop.tot', 'i.M', 'pred3')
plotdistri(Bon,c("i.M"),j, "pred3")
j <- c('tprop.E', 'tprop.EM', 'tprop.SC', 'tprop.tot', 'i.M', 'pred6')
plotdistri(Bon,c("i.M"),j, "pred6")
```

  
  
###### Just BTN draws

```{r pred36btn, echo=FALSE, include=FALSE}
# 
# j <- c('tot', 'totU','i.M',"Progression")
# plotdistri(BTN, c("i.M"), j, "Progression")
# j <- c('tot', 'totU','i.M',"pred4")
# plotdistri(BTN, c("i.M"), j, "pred4")
# j <- c('tot', 'totU','i.M',"pred8")
# plotdistri(BTN, c("i.M"), j, "pred8")
# j <- c('tot', 'totU','i.M',"pred3")
# plotdistri(BTN, c("i.M"), j, "pred3")
# j <- c('tot', 'totU','i.M',"pred6")
# plotdistri(BTN, c("i.M"), j, "pred6")
# 
# j <- c('E','EM','M','SC','numClust','i.M', "Progression")
# plotdistri(BTN,c("i.M"),j, "Progression")
# j <- c('E','EM','M','SC','numClust','i.M', "pred4")
# plotdistri(BTN,c("i.M"),j, "pred4")
# j <- c('E','EM','M','SC','numClust','i.M', "pred8")
# plotdistri(BTN, c("i.M"), j, "pred8")
# j <- c('E','EM','M','SC','numClust','i.M', "pred3")
# plotdistri(BTN,c("i.M"),j, "pred3")
# j <- c('E','EM','M','SC','numClust','i.M', "pred6")
# plotdistri(BTN, c("i.M"), j, "pred6")
# 
# j <- c('e','c','ec','ve','vc','vec','i.M',"Progression")
# plotdistri(BTN,c("i.M"),j, "Progression")
# j <- c('e','c','ec','ve','vc','vec','i.M',"pred4")
# plotdistri(BTN,c("i.M"),j, "pred4")
# j <- c('e','c','ec','ve','vc','vec','i.M',"pred8")
# plotdistri(BTN, c("i.M"), j, "pred8")
# j <- c('e','c','ec','ve','vc','vec','i.M',"pred3")
# plotdistri(BTN,c("i.M"),j, "pred3")
# j <- c('e','c','ec','ve','vc','vec','i.M',"pred6")
# plotdistri(BTN, c("i.M"), j, "pred6")
# 
# ## looking at groupings together
# j <- c('e2', 've2', 'vec', 'i.M', 'Progression')
# plotdistri(BTN,c("i.M"),j, "Progression")
# j <- c('e2', 've2', 'vec', 'i.M', 'pred4')
# plotdistri(BTN,c("i.M"),j, "pred4")
# j <- c('e2', 've2', 'vec', 'i.M', 'pred8')
# plotdistri(BTN,c("i.M"),j, "pred8")
# j <- c('e2', 've2', 'vec', 'i.M', 'pred3')
# plotdistri(BTN,c("i.M"),j, "pred3")
# j <- c('e2', 've2', 'vec', 'i.M', 'pred6')
# plotdistri(BTN,c("i.M"),j, "pred6")

j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'Progression')
plotdistri(BTN,c("i.M"),j, "Progression")
j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'pred4')
plotdistri(BTN,c("i.M"),j, "pred4")
j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'pred8')
plotdistri(BTN,c("i.M"),j, "pred8")
j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'pred3')
plotdistri(BTN,c("i.M"),j, "pred3")
j <- c('c2', 'vc2', 'vec', 've.vc', 'i.M', 'pred6')
plotdistri(BTN,c("i.M"),j, "pred6")

j <- c('CA19', 'i.M', 'Progression')
plotdistri(BTN,c("i.M"),j, "Progression")
j <- c('CA19', 'i.M', 'pred4')
plotdistri(BTN,c("i.M"),j, "pred4")
j <- c('CA19', 'i.M', 'pred8')
plotdistri(BTN,c("i.M"),j, "pred8")
j <- c('CA19', 'i.M', 'pred3')
plotdistri(BTN,c("i.M"),j, "pred3")
j <- c('CA19', 'i.M', 'pred6')
plotdistri(BTN,c("i.M"),j, "pred6")
```
  
  
  
###### Just BTN draws

```{r pred36btnlog, echo=FALSE, include=FALSE}
## log10
BTN[, c('log.tot', 'log.totU', 'log.E','log.EM','log.M','log.SC','log.e','log.c','log.ec','log.ve','log.vc','log.vec','log.e2', 'log.ve2', 'log.c2', 'log.vc2') :=
      .(
        log10(tot + 1),
        log10(totU + 1),
        log10(E + 1),
        log10(EM + 1),
        log10(M + 1),
        log10(SC + 1),
        log10(e + 1),
        log10(c + 1),
        log10(ec + 1),
        log10(ve + 1),
        log10(vc + 1),
        log10(vec + 1),
        log10(e2 + 1),
        log10(ve2 + 1),
        log10(c2 + 1),
        log10(vc2 + 1)
      )]

j <- c('log.tot', 'log.totU','i.M',"Progression")
plotdistri(BTN, c("i.M"), j, "Progression")
j <- c('log.tot', 'log.totU','i.M',"pred4")
plotdistri(BTN, c("i.M"), j, "pred4")
j <- c('log.tot', 'log.totU','i.M',"pred8")
plotdistri(BTN, c("i.M"), j, "pred8")
j <- c('log.tot', 'log.totU','i.M',"pred3")
plotdistri(BTN, c("i.M"), j, "pred3")
j <- c('log.tot', 'log.totU','i.M',"pred6")
plotdistri(BTN, c("i.M"), j, "pred6")

j <- c('log.E','log.EM','log.M','log.SC','numClust','i.M', "Progression")
plotdistri(BTN,c("i.M"),j, "Progression")
j <- c('log.E','log.EM','log.M','log.SC','numClust','i.M', "pred4")
plotdistri(BTN,c("i.M"),j, "pred4")
j <- c('log.E','log.EM','log.M','log.SC','numClust','i.M', "pred8")
plotdistri(BTN, c("i.M"), j, "pred8")
j <- c('log.E','log.EM','log.M','log.SC','numClust','i.M', "pred3")
plotdistri(BTN,c("i.M"),j, "pred3")
j <- c('log.E','log.EM','log.M','log.SC','numClust','i.M', "pred6")
plotdistri(BTN, c("i.M"), j, "pred6")

j <- c('log.e','log.c','log.ec','log.ve','log.vc','log.vec','i.M',"Progression")
plotdistri(BTN,c("i.M"),j, "Progression")
j <- c('log.e','log.c','log.ec','log.ve','log.vc','log.vec','i.M',"pred4")
plotdistri(BTN,c("i.M"),j, "pred4")
j <- c('log.e','log.c','log.ec','log.ve','log.vc','log.vec','i.M',"pred8")
plotdistri(BTN, c("i.M"), j, "pred8")
j <- c('log.e','log.c','log.ec','log.ve','log.vc','log.vec','i.M',"pred3")
plotdistri(BTN,c("i.M"),j, "pred3")
j <- c('log.e','log.c','log.ec','log.ve','log.vc','log.vec','i.M',"pred6")
plotdistri(BTN, c("i.M"), j, "pred6")

## looking at groupings together
j <- c('log.e2', 'log.ve2', 'log.vec', 'i.M', 'Progression')
plotdistri(BTN,c("i.M"),j, "Progression")
j <- c('log.e2', 'log.ve2', 'log.vec', 'i.M', 'pred4')
plotdistri(BTN,c("i.M"),j, "pred4")
j <- c('log.e2', 'log.ve2', 'log.vec', 'i.M', 'pred8')
plotdistri(BTN,c("i.M"),j, "pred8")
j <- c('log.e2', 'log.ve2', 'log.vec','i.M', 'pred3')
plotdistri(BTN,c("i.M"),j, "pred3")
j <- c('log.e2', 'log.ve2', 'log.vec','i.M', 'pred6')
plotdistri(BTN,c("i.M"),j, "pred6")

j <- c('log.c2', 'log.vc2', 'log.vec', 'i.M', 'Progression')
plotdistri(BTN,c("i.M"),j, "Progression")
j <- c('log.c2', 'log.vc2', 'log.vec', 'i.M', 'pred4')
plotdistri(BTN,c("i.M"),j, "pred4")
j <- c('log.c2', 'log.vc2', 'log.vec', 'i.M', 'pred8')
plotdistri(BTN,c("i.M"),j, "pred8")
j <- c('log.c2', 'log.vc2', 'log.vec', 'i.M', 'pred3')
plotdistri(BTN,c("i.M"),j, "pred3")
j <- c('log.c2', 'log.vc2', 'log.vec', 'i.M', 'pred6')
plotdistri(BTN,c("i.M"),j, "pred6")
```
  
###### Proportions and subs

```{r pred36distpropsub, echo=FALSE, include=FALSE}
## BTN
# j <- c('prop.tot', 'prop.totU','i.M',"pred3")
# plotdistri(B, c("i.M"), j, "pred3")
# j <- c('prop.tot', 'prop.totU','i.M',"pred6")
# plotdistri(B, c("i.M"), j, "pred6")
# 
# j <- c('prop.E','prop.EM','prop.M','prop.SC','prop.clust','i.M', "pred3")
# plotdistri(B, c("i.M"), j, "pred3")
# j <- c('prop.E','prop.EM','prop.M','prop.SC','prop.clust','i.M', "pred6")
# plotdistri(B, c("i.M"), j, "pred6")
# 
# j <- c('prop.e','prop.c','prop.ec','prop.ve','prop.vc','prop.vec','i.M',"pred3")
# plotdistri(B, c("i.M"),j, "pred3")
# j <- c('prop.e','prop.c','prop.ec','prop.ve','prop.vc','prop.vec','i.M',"pred6")
# plotdistri(B, c("i.M"), j, "pred6")

j <- c('prop.e2','prop.c2','prop.ve2','prop.vc2','prop.ve.vc','i.M',"pred3")
plotdistri(B, c("i.M"),j, "pred3")
j <- c('prop.e2','prop.c2','prop.ve2','prop.vc2','prop.ve.vc','i.M',"pred6")
plotdistri(B, c("i.M"), j, "pred6")

j <- c('prop.CA19','i.M',"pred3")
plotdistri(B, c("i.M"),j, "pred3")
j <- c('prop.CA19','i.M',"pred6")
plotdistri(B, c("i.M"), j, "pred6")


# j <- c('prop.e','prop.c','prop.ec','prop.ve','prop.vc','prop.vec','i.M',"pred3")
# plotdistri(B, c("i.M"),j, "pred3")
# j <- c('prop.e','prop.c','prop.ec','prop.ve','prop.vc','prop.vec','i.M',"pred6")
# plotdistri(B, c("i.M"), j, "pred6")

# j <- c('sub.tot', 'sub.totU','i.M',"pred3")
# plotdistri(B, c("i.M"), j, "pred3")
# j <- c('sub.tot', 'sub.totU','i.M',"pred6")
# plotdistri(B, c("i.M"), j, "pred6")
# 
# j <- c('sub.E','sub.EM','sub.M','sub.SC','sub.clust','i.M', "pred3")
# plotdistri(B, c("i.M"), j, "pred3")
# j <- c('sub.E','sub.EM','sub.M','sub.SC','sub.clust','i.M', "pred6")
# plotdistri(B, c("i.M"), j, "pred6")
# 
# j <- c('sub.e','sub.c','sub.ec','sub.ve','sub.vc','sub.vec','i.M',"pred3")
# plotdistri(B, c("i.M"),j, "pred3")
# j <- c('sub.e','sub.c','sub.ec','sub.ve','sub.vc','sub.vec','i.M',"pred6")
# plotdistri(B, c("i.M"), j, "pred6")

j <- c('sub.e2','sub.c2','sub.ve2','sub.vc2','sub.ve.vc','i.M',"pred3")
plotdistri(B, c("i.M"),j, "pred3")
j <- c('sub.e2','sub.c2','sub.ve2','sub.vc2','sub.ve.vc','i.M',"pred6")
plotdistri(B, c("i.M"), j, "pred6")

j <- c('sub.CA19','i.M',"pred3")
plotdistri(B, c("i.M"),j, "pred3")
j <- c('sub.CA19','i.M',"pred6")
plotdistri(B, c("i.M"), j, "pred6")

```
  
  
### Baseline to first draw on Firstline

```{r firstbase, echo=FALSE, include=FALSE}
bc <- B[Draw.ID != "MK06-10" & line == "First Line" &
          !is.element(Patient.ID, c("GV50", "MK105", "MK48", "MK176","MK95","MK05")),
        .SD[1], by="Patient.ID"]
j <- c('sub.tot', 'sub.totU','i.M',"pred3")
plotdistriNS(bc, c("i.M"), j, "pred3")
j <- c('sub.tot', 'sub.totU','i.M',"pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")

j <- c('sub.E','sub.EM','sub.M','sub.SC','sub.clust','i.M', "pred3")
plotdistriNS(bc, c("i.M"), j, "pred3")
j <- c('sub.E','sub.EM','sub.M','sub.SC','sub.clust','i.M', "pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")

j <- c('sub.e','sub.c','sub.ec','sub.ve','sub.vc','sub.vec','i.M',"pred3")
plotdistriNS(bc, c("i.M"),j, "pred3")
j <- c('sub.e','sub.c','sub.ec','sub.ve','sub.vc','sub.vec','i.M',"pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")

j <- c('sub.e2','sub.c2','sub.ve2','sub.vc2', 'sub.ve.vc','i.M',"pred3")
plotdistriNS(bc, c("i.M"),j, "pred3")
j <- c('sub.e2','sub.c2','sub.ve2','sub.vc2', 'sub.ve.vc','i.M',"pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")

j <- c('sub.CA19','i.M',"pred3")
plotdistriNS(bc, c("i.M"),j, "pred3")
j <- c('sub.CA19','i.M',"pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")

## Prop
j <- c('prop.tot', 'prop.totU','i.M',"pred3")
plotdistriNS(bc, c("i.M"), j, "pred3")
j <- c('prop.tot', 'prop.totU','i.M',"pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")

j <- c('prop.E','prop.EM','prop.M','prop.SC','prop.clust','i.M', "pred3")
plotdistriNS(bc, c("i.M"), j, "pred3")
j <- c('prop.E','prop.EM','prop.M','prop.SC','prop.clust','i.M', "pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")


j <- c('prop.e','prop.c','prop.ec','prop.ve','prop.vc','prop.vec','i.M',"pred3")
plotdistriNS(bc, c("i.M"),j, "pred3")
j <- c('prop.e','prop.c','prop.ec','prop.ve','prop.vc','prop.vec','i.M',"pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")

j <- c('prop.e2','prop.c2','prop.ve2','prop.vc2', 'prop.ve.vc','i.M',"pred3")
plotdistriNS(bc, c("i.M"),j, "pred3")
j <- c('prop.e2','prop.c2','prop.ve2','prop.vc2', 'prop.ve.vc','i.M',"pred6")
plotdistri(bc, c("i.M"), j, "pred6")

j <- c('prop.CA19','i.M',"pred3")
plotdistriNS(bc, c("i.M"),j, "pred3")
j <- c('prop.CA19','i.M',"pred6")
plotdistriNS(bc, c("i.M"), j, "pred6")
```

#### Baseline to first draw for First line M0 prediction of Partial Response (PR)

```{r baselinepr, echo=FALSE, include=FALSE}
bc <- B[i.M == "M0" & line == "First Line" &
          !is.element(Patient.ID, c("GV50", "MK105", "MK48", "MK176","MK95","MK05")),
        .SD[1], by="Patient.ID"]
j <- c('tot', 'totU','i.M',"PRpred3")
plotdistri(bc, c("i.M"), j, "PRpred3")
j <- c('tot', 'totU','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

j <- c('E','EM','M','SC','numClust','i.M', "PRpred3")
plotdistri(bc, c("i.M"), j, "PRpred3")
j <- c('E','EM','M','SC','numClust','i.M', "PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

j <- c('e','c','ec','ve','vc','vec','i.M',"PRpred3")
plotdistri(bc, c("i.M"),j, "PRpred3")
j <- c('e','c','ec','ve','vc','vec','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

j <- c('e2','c2','ve2','vc2', 've.vc','i.M',"PRpred3")
plotdistri(bc, c("i.M"),j, "PRpred3")
j <- c('e2','c2','ve2','vc2', 've.vc','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

## sub
j <- c('sub.tot', 'sub.totU','i.M',"PRpred3")
plotdistri(bc, c("i.M"), j, "PRpred3")
j <- c('sub.tot', 'sub.totU','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

j <- c('sub.E','sub.EM','sub.M','sub.SC','sub.clust','i.M', "PRpred3")
plotdistri(bc, c("i.M"), j, "PRpred3")
j <- c('sub.E','sub.EM','sub.M','sub.SC','sub.clust','i.M', "PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

j <- c('sub.e','sub.c','sub.ec','sub.ve','sub.vc','sub.vec','i.M',"PRpred3")
plotdistri(bc, c("i.M"),j, "PRpred3")
j <- c('sub.e','sub.c','sub.ec','sub.ve','sub.vc','sub.vec','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

j <- c('sub.e2','sub.c2','sub.ve2','sub.vc2', 'sub.ve.vc','i.M',"PRpred3")
plotdistri(bc, c("i.M"),j, "PRpred3")
j <- c('sub.e2','sub.c2','sub.ve2','sub.vc2', 'sub.ve.vc','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

## Prop
j <- c('prop.tot', 'prop.totU','i.M',"PRpred3")
plotdistri(bc, c("i.M"), j, "PRpred3")
j <- c('prop.tot', 'prop.totU','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

j <- c('prop.E','prop.EM','prop.M','prop.SC','prop.clust','i.M', "PRpred3")
plotdistri(bc, c("i.M"), j, "PRpred3")
j <- c('prop.E','prop.EM','prop.M','prop.SC','prop.clust','i.M', "PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")


j <- c('prop.e','prop.c','prop.ec','prop.ve','prop.vc','prop.vec','i.M',"PRpred3")
plotdistri(bc, c("i.M"),j, "PRpred3")
j <- c('prop.e','prop.c','prop.ec','prop.ve','prop.vc','prop.vec','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")

j <- c('prop.e2','prop.c2','prop.ve2','prop.vc2', 'prop.ve.vc','i.M',"PRpred3")
plotdistri(bc, c("i.M"),j, "PRpred3")
j <- c('prop.e2','prop.c2','prop.ve2','prop.vc2', 'prop.ve.vc','i.M',"PRpred6")
plotdistri(bc, c("i.M"), j, "PRpred6")
```


```{r ggrocs, echo=FALSE}
# Functions plots multiple 'roc' objects into one plot
# @param rocs
#   A list of 'roc' objects. Every list item has a name.
# @param breaks
#   A vector of integers representing ticks on the x- and y-axis
# @param legentTitle
#   A string which is used as legend titel
ggrocs <- function(rocs, breaks = seq(0,1,0.1), legendTitel = "Legend") {
  if (length(rocs) == 0) {
    stop("No ROC objects available in param rocs.")
  } else {
    require(plyr)
    print("Spot 1")
    # Store all sensitivities and specifivities in a data frame
    # which an be used in ggplot
    RocVals <- plyr::ldply(names(rocs), function(rocName) {
      print("Spot 2")
      if(class(rocs[[rocName]]) != "roc") {
        stop("Please provide roc object from pROC package")
      }
      print("spot 3")
      data.frame(
        fpr = rev(rocs[[rocName]]$specificities),
        tpr = rev(rocs[[rocName]]$sensitivities),
        names = rep(rocName, length(rocs[[rocName]]$sensitivities)),
        stringAsFactors = T
      )
    })
    print(RocVals)
    
    aucAvg <- mean(sapply(rocs, "[[", "auc"))

    rocPlot <- ggplot(RocVals, aes(x = fpr, y = tpr, colour = names)) +
      geom_segment(aes(x = 0, y = 1, xend = 1,yend = 0), alpha = 0.5, colour = "gray") + 
      geom_step() +
      scale_x_reverse(name = "False Positive Rate (1 - Specificity)",limits = c(1,0), breaks = breaks) + 
      scale_y_continuous(name = "True Positive Rate (Sensitivity)", limits = c(0,1), breaks = breaks) +
      theme_bw() + 
      coord_equal() + 
      annotate("text", x = 0.1, y = 0.1, vjust = 0, label = paste("AUC =",sprintf("%.3f",aucAvg))) +
      guides(colour = guide_legend(legendTitel)) +
      theme(axis.ticks = element_line(color = "grey80"))
    
    rocPlot
  }
}
```


#### Evaluation of data through ROCs

```{r samp36ROC, echo=FALSE}
ROC <- function(dt, pred, vars, title){
  cols <- hue_pal()(2*length(vars))
  for (i in 1:length(vars)) {
    plot.roc(dt[i.M == "M0", get(pred)], dt[i.M == "M0", get(vars[i])], col = cols[i], main=paste(title, "variable:", vars[i]), percent=TRUE, ci=TRUE, of="thresholds",thresholds="best", print.thres="best", asp=NA, print.auc=TRUE)
    lines.roc(dt[i.M == "M1", get(pred)], dt[i.M == "M1", get(vars[i])], col = cols[i + length(vars)], main=paste(title, "variable:", vars[i]), percent=TRUE, ci=TRUE, of="thresholds",thresholds="best", print.thres="best", asp=NA,print.auc=TRUE,add=TRUE)
  legend("bottomright", legend=c("M0", "M1"), col=c(cols[i], cols[i + length(vars)]), lwd=2) 
  }
}
ROCc <- function(dt, pred, vars, title){
  cols <- hue_pal()(2*length(vars))
  for (i in 1:length(vars)) {
    plot.roc(dt[, get(pred)], dt[, get(vars[i])], col = cols[i], main=paste(title, "variable:", vars[i]), percent=TRUE, ci=TRUE, of="thresholds",thresholds="best", print.thres="best")
  legend("bottomright", legend=c("M0 + M1"), col=c(cols[i], cols[i + length(vars)]), lwd=2) 
  }
}
ROC2 <- function(dt, pred, vars, title){
  cols <- hue_pal()(2*length(vars))
  for (i in 1:length(vars)) {
    p <- plot.roc(dt[i.M == "M0", get(pred)], dt[i.M == "M0", get(vars[i])], col = cols[i], main=paste(title, "variable:", vars[i]), percent=TRUE, ci=TRUE, of="thresholds",thresholds="best", print.thres="best", asp=NA, print.auc=TRUE)
    p <- plot.roc(dt[i.M == "M1", get(pred)], dt[i.M == "M1", get(vars[i])], col = cols[i + length(vars)], main=paste(title, "variable:", vars[i]), percent=TRUE, ci=TRUE, of="thresholds",thresholds="best", print.thres="best", asp=NA,print.auc=TRUE, print.auc.y=40, add=TRUE)
  p <- legend("bottomright", legend=c("M0", "M1"), col=c(cols[i], cols[i + length(vars)]), lwd=2) 
  }
}
# ROC <- function(dt, x, preds){
#   rocs <- list()
#   for(i in 1:length(x)){
#     rocs[[i]] <- roc(dt[, get(preds)], dt[, get(x)])
#   }
#   return(rocs)
# }
# variable sets.
vars <- c("e","c","ec","ve","ve2","vc","vc2",'ve.vc',"vec","E","EM","M","SC","U","tot","totU", "numClust","CA19")
varsMean <- c("nMean.e","nMean.c","nMean.ec","nMean.ve","nMean.vc","nMean.vec","nMean.E","nMean.EM","nMean.M","nMean.SC","nMean.U","nMean.tot","nMean.totU")
varsMed <- c("nMedian.e","nMedian.c","nMedian.ec","nMedian.ve","nMedian.vc","nMedian.vec","nMedian.E","nMedian.EM","nMedian.M","nMedian.SC","nMedian.U","nMedian.tot","nMedian.totU")
varsP <- c("prop.e","prop.e2","prop.c","prop.c2","prop.ec","prop.ve","prop.ve2","prop.vc","prop.vc2",'prop.ve.vc',"prop.vec","prop.E","prop.EM","prop.M","prop.SC","prop.U","prop.tot","prop.totU", "prop.clust","prop.CA19")
varsS <- c("sub.e","sub.e2","sub.c","sub.c2","sub.ec","sub.ve","sub.ve2","sub.vc","sub.vc2",'sub.ve.vc',"sub.vec","sub.E","sub.EM","sub.M","sub.SC","sub.U","sub.tot","sub.totU", "sub.clust","sub.CA19")
```

##### For all patients

Don't necessarily expect to find anything here. Purpose is "just in case"
Split between M1 and M0

###### Single Sample as point

Takes a single draw and compares within time constraint

```{r samp36all, echo=FALSE, include=FALSE}
# basic counts
# vars <- c('ve.vc',"CA19")
ROC(A, "pred3", vars, "ROC for prediction with 3 month time window, ")
ROC(A, "pred6", vars, "ROC for prediction with 6 month time window, ")

# nMean
# ROC(A, "pred3", varsMean, "ROC for prediction with 3 month time window, ")
# ROC(A, "pred6", varsMean, "ROC for prediction with 6 month time window, ")
# 
# # nMedian
# ROC(A, "pred3", varsMed, "ROC for prediction with 3 month time window, ")
# ROC(A, "pred6", varsMed, "ROC for prediction with 6 month time window, ")
```

###### Based off delta

Takes in delta between points

```{r samp36alldelta, echo=FALSE, include=FALSE}

```

##### For treatment Naives

Don't necessarily expect to find anything here. Purpose is "just in case"

###### Single Sample as Point

```{r samp36btn, echo=FALSE, include=FALSE}
# baseline
#varsP <- c('prop.EM',"prop.E","prop.tot")
#varsS <- c('sub.EM',"sub.E", "sub.tot")
# Bt <- B[!is.element(Draw.ID, bc[,Draw.ID])]
neoadj <- c("MK07-2","MK100-2","MK115-2","MK145-2","MK40-2","MK51-2","MK55-2","MK81-2","MK84-2","MK98-2")
Bt <- B[!is.element(Draw.ID, neoadj)]
ROC(Bt, "pred3", varsP, "ROC for prediction with 3 month time window, ")
# ROC(Bt, "pred6", varsP, "ROC for prediction with 6 month time window, ")

ROC(Bt, "pred3", varsS, "ROC for prediction with 3 month time window, ")
# ROC(Bt, "pred6", varsS, "ROC for prediction with 6 month time window, ")
```


##### t/p by line of therapy, no baselines.
```{r roctp, echo=FALSE}
vars <- c("tve","tvc","tvec","tE","tEM","tM","tSC","ttot","pve","pvc","pvec","pE","pEM","pM","pSC","ptot")
ROC(Aline, "pred3", vars, "ROC for prediction with 3 month time window, ")
```


##### Cut offs at 3 months KM

```{r KM3months, echo=FALSE}
# Proportions
# B[, c('prop.e','prop.e2','prop.c','prop.c2','prop.ec','prop.ve','prop.vc','prop.ve2','prop.vc2','prop.vec',"prop.E","prop.EM","prop.M","prop.SC","prop.tot","prop.totU", "prop.clust") := 
B[, c('prop.c2co',"prop.Eco","prop.EMco","prop.EMco2","prop.EMco3","prop.totco","prop.totco2") :=
         .(
           ifelse(prop.c2 >= 0.8, 1, 0),
           ifelse(prop.E >= 0.9, 1, 0),
           ifelse(prop.EM >= 0.6, 1, 0),
           ifelse(prop.EM >= 1.3, 1, 0),
           ifelse(prop.EM >= 1.4, 1, 0),
           ifelse(prop.tot >= 1, 1, 0),
           ifelse(prop.tot >= 1.1, 1, 0)
         )]

# Subtractions
# B[, c('sub.e','sub.c','sub.ec','sub.ve','sub.vc','sub.vec',"sub.E","sub.EM","sub.M","sub.SC","sub.tot","sub.totU", "sub.clust") := 
B[, c('sub.c2co','sub.c2co2',"sub.Eco","sub.EMco","sub.EMco2","sub.totco") := 
         .(
           ifelse(sub.c2 >= 0.8, 1, 0),
           ifelse(sub.c2 >= -0.5, 1, 0),
           ifelse(sub.E >= 0.8, 1, 0),
           ifelse(sub.EM >= -0.5, 1, 0),
           ifelse(sub.EM >= -6.5, 1, 0),
           ifelse(sub.tot >= -12, 1, 0)
         )]
BTN[, "EMco" := ifelse(EM >= 5 , 1, 0)]
BTN[, "clustco" := ifelse(numClust >= 0.5 , 1, 0)]
BTN[, c("cfKRAS.co","exoKRAS.co") := .(
        ifelse(cfKRAS > 0, 1, ifelse(is.na(cfKRAS),cfKRAS, 0)),
        ifelse(exoKRAS > 5, 1, ifelse(is.na(cfKRAS),exoKRAS, 0))
)]

Aline[, c("tvc.co","tEM.co","ttot.co","pvc.co","pEM.co","ptot.co") := 
         .(
           ifelse(tvc >= 2.5, 1, 0),
           ifelse(tEM >= 5, 1, 0),
           ifelse(ttot >= 9.5, 1, 0),
           ifelse(pvc >= 2, 1, 0),
           ifelse(pEM >= 2.1, 1, 0),
           ifelse(ptot >= 1.7, 1, 0)
         )]

library(survival)
library(stargazer)
create2by2 <- function(column, pred){
  print(column)
  a <- B[i.M == "M0" & eval(as.name(column)) == 1 & eval(as.name(pred)) == 1, .N]
  b <- B[i.M == "M0" & eval(as.name(column)) == 0 & eval(as.name(pred)) == 1, .N]
  c <- B[i.M == "M0" & eval(as.name(column)) == 1 & eval(as.name(pred)) == 0, .N]
  d <- B[i.M == "M0" & eval(as.name(column)) == 0 & eval(as.name(pred)) == 0, .N]
  m <- matrix(c(a,b,c,d),ncol = 2)
  stargazer(m, type="text")
  fisher.test(m)
}
subs <- c('sub.c2co','sub.c2co2',"sub.Eco","sub.EMco","sub.totco")
props <- c('prop.c2co',"prop.Eco","prop.EMco","prop.totco")
lapply(subs, pred="pred3", create2by2)
lapply(props, pred="pred3", create2by2)
```


#### Training and Validation Sets

```{r trainval, echo=FALSE}
create2by2dt <- function(column, pred, dt){
  print(column)
  a <- dt[i.M == "M0" & eval(as.name(column)) == 1 & eval(as.name(pred)) == 1, .N]
  b <- dt[i.M == "M0" & eval(as.name(column)) == 0 & eval(as.name(pred)) == 1, .N]
  c <- dt[i.M == "M0" & eval(as.name(column)) == 1 & eval(as.name(pred)) == 0, .N]
  d <- dt[i.M == "M0" & eval(as.name(column)) == 0 & eval(as.name(pred)) == 0, .N]
  m <- matrix(c(a,b,c,d),ncol = 2)
  stargazer(m, type="text")
  fisher.test(m)
}
# subs <- c('sub.c2co','sub.c2co2',"sub.Eco","sub.EMco","sub.totco")
# props <- c('prop.c2co',"prop.Eco","prop.EMco2","prop.totco")
# lapply(subs, pred="pred3", dt=B[!is.element(Draw.ID, bc[,Draw.ID])], create2by2dt)
# lapply(props, pred="pred3", dt=B[!is.element(Draw.ID, bc[,Draw.ID])], create2by2dt)
# subs <- c("sub.E","sub.EMco2","sub.totco")
# props <- c("prop.E","prop.EMco3","prop.totco2")
# lapply(subs, pred="pred3", dt=B[is.element(Draw.ID, neoadj)], create2by2dt)
# lapply(props, pred="pred3", dt=B[is.element(Draw.ID, neoadj)], create2by2dt)


lapply(c("tvc.co","tEM.co","ttot.co","pvc.co","pEM.co","ptot.co"), pred="pred3", dt=Aline, create2by2dt)
```


##### Logistic Regressions

```{r logregpred3}
# lr <- B
# lr[, pred3 := as.numeric(pred3) - 1]
# lapply(subs, function(X){
#   ggplot(lr, aes(x = X, y=pred3)) + geom_point() + geom_smooth(method = "glm", method.args=list(family = "binomial"), se=FALSE)
# })

```

## Patient Based

```{r pat, echo=FALSE}

```
