---
title: "Unresectable Cox Reg Data"
author: "Jonathan"
date: "November 28, 2017"
output:
  pdf_document: default
  html_document:
    df_print: paged
  tex_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(survminer)
library(ggplot2)
library(ggpubr)
library(stargazer)
df = read.csv('Unresectable/CoxReg/genericfilename.csv', sep=',', na.strings = c('',' '))
met = read.csv('Unresectable/CoxReg/genericfilename1.csv', sep=',', na.strings = c('',' '))
la = read.csv('Unresectable/CoxReg/genericfilename2.csv', sep=',', na.strings = c('',' '))
resect = read.csv('Resectable/Data/genericfilename.csv', sep=',', na.strings = c('',' '))
## Added log scales to some fields
## CA19
df$logCA19 <- log2(df$CA19.9)
## exoDNA
##   Base 2
df$log2exo <- log2(df$exoDNA + 1)
df$log2exong <- log2(df$exoDNA.mut.ng.ml+1)
df$log2exotot <- log2(df$total.DNA.yield.exo+1)
df$log2exogen <- log2(df$genomic.equivalents.exo+1)
### Base 10
df$log10exo <- log10(df$exoDNA + 1)
df$log10exong <- log10(df$exoDNA.mut.ng.ml+1)
df$log10exotot <- log10(df$total.DNA.yield.exo+1)
df$log10exogen <- log10(df$genomic.equivalents.exo+1)
## cfDNA
##   Base 2
df$log2cf <- log2(df$cfDNA + 1)
df$log2cfng <- log2(df$cfDNA.mut.ng.ml+1)
df$log2cftot <- log2(df$total.DNA.yield.cf+1)
df$log2cfgen <- log2(df$genomic.equivalents.cf+1)
##   Base 10
df$log10cf <- log10(df$cfDNA + 1)
df$log10cfng <- log10(df$cfDNA.mut.ng.ml+1)
df$log10cftot <- log10(df$total.DNA.yield.cf+1)
df$log10cfgen <- log10(df$genomic.equivalents.cf+1)

## for just met
met$logCA19 <- log2(met$CA19.9)
met$CA19p1 <- met$CA19.9+1
## exoDNA
## + 1
met$exop1 <- met$exoDNA + 1
##   Base 2
met$log2exo <- log2(met$exoDNA + 1)
met$log2exong <- log2(met$exoDNA.mut.ng.ml+1)
met$log2exotot <- log2(met$total.DNA.yield.exo+1)
met$log2exogen <- log2(met$genomic.equivalents.exo+1)

## base 10
met$log10exo <- log10(met$exoDNA + 1)

## cfDNA
## + 1
met$cfp1 <- met$cfDNA + 1
##   Base 2
met$log2cf <- log2(met$cfDNA + 1)
met$log2cfng <- log2(met$cfDNA.mut.ng.ml+1)
met$log2cftot <- log2(met$total.DNA.yield.cf+1)
met$log2cfgen <- log2(met$genomic.equivalents.cf+1)

## base 10
met$log10cf <- log10(met$cfDNA + 1)


## la
la$logCA19 <- log2(la$CA19.9)
la$log2exo <- log2(la$exoDNA + 1)
la$log2exong <- log2(la$exoDNA.mut.ng.ml+1)
la$log2exotot <- log2(la$total.DNA.yield.exo+1)
la$log2exogen <- log2(la$genomic.equivalents.exo+1)

la$log2cf <- log2(la$cfDNA + 1)
la$log2cfng <- log2(la$cfDNA.mut.ng.ml+1)
la$log2cftot <- log2(la$total.DNA.yield.cf+1)
la$log2cfgen <- log2(la$genomic.equivalents.cf+1)

## resectables
resect$exop1 <- resect$exoDNA + 1
resect$cfp1 <- resect$cfDNA + 1
resect$CA19p1 <- resect$cfDNA + 1
```

# Univariate

Univariate cox regressions for most variables found in our demographics tables

## Both

### PFS

```{r univariatePFS, echo=FALSE}
vars <- c('exoDNA', 'exo5', 'exo1', 'exo', 'log2exo','total.DNA.yield.exo', 'exoDNA.mut.ng.ml', 'genomic.equivalents.exo', 'log2cf', 'cfDNA', 'cf1', 'cf', 'total.DNA.yield.cf', 'cfDNA.mut.ng.ml', 'genomic.equivalents.cf', 'exoorcf', 'exoandcf', 'Age', 'gender', 'logCA19', 'ecog', 'Regimen', 'SLD', 'metLivervAll', 'CA19.300', 'Num.Met')
univ_formulas <- sapply(vars, function(x) as.formula(paste('Surv(PFS, Progression)~', x)))
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = df)})
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=3)
                          wald.test<-signif(x$wald["test"], digits=3)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 3)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],3)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = TRUE))
## as.data.frame(res)
stargazer(res, type = 'text', title='Univariate Cox Regression (PFS)')
summary(coxph(Surv(PFS, Progression)~factor(factorMet), data = df))
summary(coxph(Surv(PFS, Progression)~factor(ecog), data = df))
```

### OS

```{r univariateOS, echo=FALSE}
univ_formulas <- sapply(vars, function(x) as.formula(paste('Surv(OS, vital.status)~', x)))
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = df)})
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = TRUE))
## as.data.frame(res)
stargazer(res, type = 'text', title='Univariate Cox Regression (OS)')
summary(coxph(Surv(OS, vital.status)~factor(factorMet), data = df))
summary(coxph(Surv(OS, vital.status)~factor(ecog), data = df))
```

## Met Cohort Only

Stage IV only as LA has some confounding problems.

### PFS

```{r metcohortuniPFS, echo=FALSE}
univ_formulas <- sapply(vars, function(x) as.formula(paste('Surv(PFS, Progression)~', x)))
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = met)})
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = TRUE))
## as.data.frame(res)
stargazer(res, type = 'text', title='Univariate Cox Regression (PFS)')
summary(coxph(Surv(PFS, Progression)~factor(factorMet), data = met))
summary(coxph(Surv(PFS, Progression)~factor(ecog), data = met))
```

### OS

```{r metcohortuniOS, echo=FALSE}
univ_formulas <- sapply(vars, function(x) as.formula(paste('Surv(OS, vital.status)~', x)))
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = met)})
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = TRUE))
## as.data.frame(res)
stargazer(res, type = 'text', title='Univariate Cox Regression (OS)')
summary(coxph(Surv(OS, vital.status)~factor(factorMet), data = met))
summary(coxph(Surv(OS, vital.status)~factor(ecog), data = met))
```

## LA Cohort Only

Univariate analysis of LA cohort.

### PFS

```{r lacohortuniPFS, echo=FALSE}
vars <- c('exoDNA', 'exo5', 'exo1', 'exo', 'log2exo','total.DNA.yield.exo', 'exoDNA.mut.ng.ml', 'genomic.equivalents.exo', 'log2cf', 'cfDNA', 'cf1', 'cf', 'total.DNA.yield.cf', 'cfDNA.mut.ng.ml', 'genomic.equivalents.cf', 'exoorcf', 'exoandcf', 'Age', 'gender', 'logCA19', 'ecog', 'Regimen', 'SLD', 'CA19.300')
univ_formulas <- sapply(vars, function(x) as.formula(paste('Surv(PFS, Progression)~', x)))
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = la)})
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = TRUE))
## as.data.frame(res)
stargazer(res, type = 'text', title='Univariate Cox Regression (PFS)')
```
```{r lacohortuniOS, echo=FALSE}
vars <- c('exoDNA', 'exo5', 'exo1', 'exo', 'log2exo','total.DNA.yield.exo', 'exoDNA.mut.ng.ml', 'genomic.equivalents.exo', 'log2cf', 'cfDNA', 'cf1', 'cf', 'total.DNA.yield.cf', 'cfDNA.mut.ng.ml', 'genomic.equivalents.cf', 'exoorcf', 'exoandcf', 'Age', 'gender', 'logCA19', 'ecog', 'Regimen', 'SLD', 'CA19.300')
univ_formulas <- sapply(vars, function(x) as.formula(paste('Surv(OS, vital.status)~', x)))
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = la)})
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = TRUE))
## as.data.frame(res)
stargazer(res, type = 'text', title='Univariate Cox Regression (OS)')
```

# Multivariate

Multivariate Cox Regression (PFS). Covariates are determined by previous significance in other publications and from p values under ~.25

## Both

```{r multivariate, echo=FALSE}
covars <- c('gender', 'factor(ecog)', 'Regimen', 'factor(factorMet)', 'CA19.300')
covars.col <- paste(covars, collapse="+")
form <- as.formula(paste("Surv(PFS, Progression) ~ exo5+cf+",covars.col,sep=''))
pfs.cox <- coxph(form, data=df )
# pfsexo.cox <- coxph(Surv(PFS, Progression) ~ exo5 + SLD + factor(factorMet)  + CA19.300 + Regimen + gender + ecog, data=df )
# pfscf.cox <- coxph(Surv(PFS, Progression) ~ cf + SLD + factor(factorMet)  + CA19.300 + Regimen +gender + ecog, data=df )
# form <- as.formula(paste("Surv(PFS, Progression) ~ exoorcf+",covars.col,sep=''))
# pfsor.cox <- coxph(form, data=df )
# form <- as.formula(paste("Surv(PFS, Progression) ~ exoandcf+",covars.col,sep=''))
# pfsand.cox <- coxph(form, data=df )

covars <- c('ecog', 'Regimen', 'factor(factorMet)', 'CA19.300', 'SLD')
covars.col <- paste(covars, collapse="+")
form <- as.formula(paste("Surv(OS, vital.status) ~ exo5+cf+",covars.col,sep=''))
os.cox <- coxph(form, data=df )
# osexo.cox <- coxph(Surv(OS, vital.status) ~ exo5 + SLD + factor(factorMet) +CA19.300 +  gender + Regimen, data=df )
# oscf.cox <- coxph(Surv(OS, vital.status) ~ cf + SLD + factor(factorMet)  + CA19.300 + gender + Regimen, data=df )
# form <- as.formula(paste("Surv(OS, vital.status) ~ exoorcf+",covars.col,sep=''))
# osor.cox <- coxph(form, data=df )
# form <- as.formula(paste("Surv(OS, vital.status) ~ exoandcf+",covars.col,sep=''))
#osand.cox <- coxph(form, data=df )

stargazer(pfs.cox, os.cox, type='text', title='Multivariate Cox Regresssion Analysis')
#stargazer(pfsor.cox, pfsand.cox, osor.cox, osand.cox, type='text', title='Multivariate Cox Regression (and/or)')
```

### Summarys

Summary's of each regression

#### PFS


```{r pfs.cox}
pfs.cox
```

exoDNA or cfDNA
```{r pfsor.cox}
# pfsor.cox
```

exoDNA and cfDNA
```{r pfsand.cox}
# pfsand.cox
```

#### OS

```{r os.cox}
os.cox
```

exoDNA or cfDNA
```{r osor.cox}
# osor.cox
```

exoDNA and cfDNA
```{r osand.cox}
# osand.cox
```

### Residuals
```{r bothresid, echo=FALSE}
resid.pfs.cox <- cox.zph(pfs.cox)
# resid.pfsor.cox <- cox.zph(pfsor.cox)
#resid.pfsand.cox <- cox.zph(pfsand.cox)
resid.os.cox <- cox.zph(os.cox)
# resid.osor.cox <- cox.zph(osor.cox)
# resid.osand.cox <- cox.zph(osand.cox)
```
#### PFS

```{r resid.pfs.cox}
ggcoxzph(resid.pfs.cox)
resid.pfs.cox
```

exoDNA or cfDNA
```{r resid.pfsor.cox}
# ggcoxzph(resid.pfsor.cox)
# resid.pfsor.cox
```

exoDNA or cfDNA
```{r resid.pfsand.cox}
# ggcoxzph(resid.pfsand.cox)
# resid.pfsand.cox
```

#### OS

```{r resid.os.cox}
ggcoxzph(resid.os.cox)
resid.os.cox
```

exoDNA or cfDNA
```{r resid.osor.cox}
# ggcoxzph(resid.osor.cox)
# resid.osor.cox
```

exoDNA and cfDNA
```{r resid.osand.cox}
# ggcoxzph(resid.osand.cox)
# resid.osand.cox
```
## Multivariate for Met Cohort only

LA population is too small for multivariate analysis

```{r metmulti, echo=FALSE}
covars <- c('gender', 'factor(ecog)', 'Regimen', 'factor(factorMet)', 'CA19.300')
covars.col <- paste(covars, collapse="+")
form <- as.formula(paste("Surv(PFS, Progression) ~ exo5+cf+",covars.col,sep=''))
metpfs.cox <- coxph(form, data=met )
# metpfsexo.cox <- coxph(Surv(PFS, Progression) ~ exo5 + SLD + factor(factorMet)  + CA19.300 + Regimen + gender, data=met )
# metpfscf.cox <- coxph(Surv(PFS, Progression) ~ cf + SLD + factor(factorMet)  + CA19.300 + Regimen +gender, data=met )
# form <- as.formula(paste("Surv(PFS, Progression) ~ exoorcf+",covars.col,sep=''))
# metpfsor.cox <- coxph(form, data=met )
# form <- as.formula(paste("Surv(PFS, Progression) ~ exoandcf+",covars.col,sep=''))
# metpfsand.cox <- coxph(form, data=met )

covars <- c('ecog', 'Regimen', 'factor(factorMet)', 'CA19.300')
covars.col <- paste(covars, collapse="+")
metos.cox <- coxph(Surv(OS, vital.status) ~ exo5 + cf + SLD + factor(factorMet) + CA19.300 + gender + Regimen, data=met )
#metosexo.cox <- coxph(Surv(OS, vital.status) ~ exo5 + SLD + factor(factorMet) +CA19.300 +  gender + Regimen, data=met )
#metoscf.cox <- coxph(Surv(OS, vital.status) ~ cf + SLD + factor(factorMet)  + CA19.300 + gender + Regimen, data=met )
# metosor.cox <- coxph(Surv(OS, vital.status) ~ exoorcf + SLD + factor(factorMet)  + CA19.300 + gender + Regimen, data=met )
# metosand.cox <- coxph(Surv(OS, vital.status) ~ exoandcf + SLD + factor(factorMet) + CA19.300 + gender + Regimen, data=met )

stargazer(metpfs.cox, metos.cox, type='text', title='Multivariate Cox Regresssion Analysis')
# stargazer(metpfsor.cox, metpfsand.cox, metosor.cox, metosand.cox, type='text', title='Multivariate Cox Regression (and/or)')
```

### Summarys

Summary's of each regression

#### PFS


```{r metpfs.cox}
summary(metpfs.cox)
```

exoDNA or cfDNA
```{r metpfsor.cox}
# summary(metpfsor.cox)
```

exoDNA and cfDNA
```{r metpfsand.cox}
# summary(metpfsand.cox)
```

#### OS

```{r metos.cox}
summary(metos.cox)
```

exoDNA or cfDNA
```{r metosor.cox}
# summary(metosor.cox)
```

exoDNA and cfDNA
```{r metosand.cox}
# summary(metosand.cox)
```

### Residuals
```{r metresid, echo=FALSE}
resid.metpfs.cox <- cox.zph(metpfs.cox)
#resid.metpfsor.cox <- cox.zph(metpfsor.cox)
# resid.metpfsand.cox <- cox.zph(metpfsand.cox)
resid.metos.cox <- cox.zph(metos.cox)
# resid.metosor.cox <- cox.zph(metosor.cox)
# resid.metosand.cox <- cox.zph(metosand.cox)
```
#### PFS

```{r resid.metpfs.cox}
ggcoxzph(resid.metpfs.cox)
resid.metpfs.cox
```

exoDNA or cfDNA
```{r resid.metpfsor.cox}
# ggcoxzph(resid.metpfsor.cox)
# resid.metpfsor.cox
```

exoDNA or cfDNA
```{r resid.metpfsand.cox}
# ggcoxzph(resid.metpfsand.cox)
# resid.metpfsand.cox
```

#### OS

```{r resid.metos.cox}
ggcoxzph(resid.metos.cox)
resid.metos.cox
```

exoDNA or cfDNA
```{r resid.metosor.cox}
# ggcoxzph(resid.metosor.cox)
# resid.metosor.cox
```

exoDNA or cfDNA
```{r resid.metosand.cox}
# ggcoxzph(resid.metosand.cox)
# resid.metosand.cox
```

# Plots

```{r plotOptions, echo=FALSE}
axii.text <- element_text(face = "bold", size = 20)

```

## Kaplan-Meier
Kaplan meier's are met only
```{r kmsetup, echo=FALSE}
### PFS
pfs.km <- survfit(Surv(PFS, Progression) ~ 1, data=met)
pfs.exo5.km <- survfit(Surv(PFS, Progression) ~ exo5, data=met)
pfs.cf.km <- survfit(Surv(PFS, Progression) ~ cf, data=met)
# pfs.exoorcf.km <- survfit(Surv(PFS, Progression) ~ exoorcf, data=met)
# pfs.exoandcf.km <- survfit(Surv(PFS, Progression) ~ exoandcf, data=met)
### OS
os.km <- survfit(Surv(OS, vital.status) ~ 1, data=met)
os.exo5.km <- survfit(Surv(OS, vital.status) ~ exo5, data=met)
os.cf.km <- survfit(Surv(PFS, vital.status) ~ cf, data=met)
# os.exoorcf.km <- survfit(Surv(PFS, vital.status) ~ exoorcf, data=met)
# os.exoandcf.km <- survfit(Surv(PFS, vital.status) ~ exoandcf, data=met)
```
### Plain Jane KM for PFS and OS

PFS
```{r pfs.km}
ggsurvplot(pfs.km, palette="#2E9FDF")
```

OS
```{r os.km}
ggsurvplot(os.km, palette="#2E9FDF")
```

### ExoDNA MAF > 5%

PFS
```{r pfs.exo5.km}
ggsurvplot(pfs.exo5.km)
```

OS
```{r os.exo5.km}
ggsurvplot(os.exo5.km)
```

### CfDNA detection
PFS
```{r pfs.cf.km}
ggsurvplot(pfs.cf.km)
```
OS
```{r os.cf.km}
ggsurvplot(os.cf.km)
```

### ExoDNA or cfDNA
PFS
```{r pfs.exoorcf.km}
# ggsurvplot(pfs.exoorcf.km)
```
OS
```{r os.exoorcf.km}
# ggsurvplot(os.exoorcf.km)
```

### ExoDNA and cfDNA
PFS
```{r pfs.exoandcf.km}
# ggsurvplot(pfs.exoandcf.km)
```
OS
```{r os.exoandcf.km}
# ggsurvplot(os.exoandcf.km)
```

## Non Survival Related

These include GLM and Boxplots

### Both Met and LA

#### GLM 

Covers mostly the supplementary tables that we may use. These graphs are both LA and MET cohorts including all metastasis.

```{r plotsGLMboth, echo=FALSE}
## GLM
## exoDNA
SLDvexo <- ggplot(df, aes(x=SLD, y=log2exo)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm') + theme_bw(base_size = 14) + labs(x='SLD (mm)', y=expression(log[2](exoDNA~MAF + 1))) + theme(legend.text=element_text(size=9))
SLDvexong <- ggplot(df, aes(x=SLD, y=log2exong)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm') + theme_bw(base_size = 14) + labs(x='SLD (mm)', y=expression(log[2](exoDNA~ng/mL + 1))) + theme(legend.text=element_text(size=9))
plot(SLDvexo)
plot(SLDvexong)

## CFDNA
SLDvcf <- ggplot(df, aes(x=SLD, y=log2cf)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm') + theme_bw(base_size=14) + labs(x='SLD (mm)', y=expression(log[2](cfDNA~MAF + 1))) + theme(legend.text=element_text(size=9))
SLDvcfng <- ggplot(df, aes(x=SLD, y=log2cfng)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm')+ theme_bw(base_size=14) + labs(x='SLD (mm)', y=expression(log[2](cfDNA~ng/mL + 1))) + theme(legend.text=element_text(size=9))
plot(SLDvcf)
plot(SLDvcfng)

## vs one another
exovcf <- ggplot(df, aes(x=log2exo, y=log2cf)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm')+ theme_bw(base_size=14) + labs(x=expression(log[2](exoDNA~MAF + 1)), y=expression(log[2](cfDNA~MAF + 1))) + theme(legend.text=element_text(size=9))
plot(exovcf)
```

#### Box/Jitter Both

```{r plotsJitterBoxBoth, echo=FALSE}
## Box/Jitter
## Met vs 
Metvexo <- ggplot(df, aes(x=Metastasis, y=log2exo, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=.01) + geom_jitter()+ theme_bw(base_size=14) + guides(color = FALSE, shape=FALSE) + labs(y=expression(log[2](exoDNA~MAF +1)))
Metvcf <- ggplot(df, aes(x=Metastasis, y=log2cf, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=0.01) + geom_jitter()+ theme_bw(base_size = 14)+ guides(color = FALSE, shape=FALSE) + labs(y=expression(log[2](cfDNA~MAF +1)))
MetvSLD <- ggplot(df, aes(x=Metastasis, y=SLD, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=.01) + geom_jitter()+ theme_bw(base_size=14) + guides(color = FALSE, shape=FALSE) + labs(y='SLD (mm)')
plot(Metvexo)
plot(Metvcf)
plot(MetvSLD)

## LA/Met
stagevexo <- ggplot(df, aes(x=clinical.stage, y=log2exo)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter( aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size=14) + labs(y=expression(log[2](exoDNA~MAF + 1)), x='Clinical Stage')+ theme(legend.text=element_text(size=9))
plot(stagevexo)
stagevexong <- ggplot(df, aes(x=clinical.stage, y=log2exong)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis))+ theme_bw(base_size=14) + labs(y=expression(log[2](exoDNA~ng/mL + 1)), x='Clinical Stage')+ theme(legend.text=element_text(size=9))
plot(stagevexong)
stagevcf <- ggplot(df, aes(x=clinical.stage, y=log2cf)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis))+ theme_bw(base_size=14) + labs(y=expression(log[2](cfDNA~MAF + 1)), x='Clinical Stage')
plot(stagevcf)+ theme(legend.text=element_text(size=9))
stagevcfng <- ggplot(df, aes(x=clinical.stage, y=log2cfng)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis))+ theme_bw(base_size=14) + labs(y=expression(log[2](cfDNA~ng/mL + 1)), x='Clinical Stage')+ theme(legend.text=element_text(size=9))
plot(stagevcfng)

## Prog/none
progvexo <- ggplot(df, aes(x=factor(Progression), y=log2exo)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size = 14) + labs(y=expression(log[2](exoDNA~MAF + 1)), x='Progression') + scale_x_discrete(labels=c("0" = "No Progression", "1" = "Progression"))+ theme(legend.text=element_text(size=9))
plot(progvexo)
progvexong <- ggplot(df, aes(x=factor(Progression), y=log2exong)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size = 14) + labs(y=expression(log[2](exoDNA~ng/mL + 1)), x='Progression') + scale_x_discrete(labels=c("0" = "No Progression", "1" = "Progression"))+ theme(legend.text=element_text(size=9))
plot(progvexong)
progvcf <- ggplot(df, aes(x=factor(Progression), y=log2cf)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size = 14) + labs(y=expression(log[2](cfDNA~MAF + 1)), x='Progression') + scale_x_discrete(labels=c("0" = "No Progression", "1" = "Progression"))+ theme(legend.text=element_text(size=9))
plot(progvcf)
progvcfng <- ggplot(df, aes(x=factor(Progression), y=log2cfng)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size = 14) + labs(y=expression(log[2](exoDNA~ng/mL + 1)), x='Progression') + scale_x_discrete(labels=c("0" = "No Progression", "1" = "Progression")) + theme(legend.text=element_text(size=9))
plot(progvcfng)

```

### Just Met

#### GLM

```{r plotsGLMmet, echo=FALSE}
## GLM
## exoDNA
SLDvexo <- ggplot(met, aes(x=SLD, y=exoDNA)) + geom_point(aes(color=Metastasis, shape=Metastasis), size=2.5) + stat_smooth(method='glm') + theme_bw(base_size = 25) + labs(x='SLD (mm)', y=expression(exo*italic(KRAS)~MAF)) + theme(legend.text=element_text(size=15)) + annotate("text", x=160, y=50, label="f(x) = 1.336 + 0.07376x \n p = 0.0353", size = 10)
SLDvexong <- ggplot(met, aes(x=SLD, y=log2exong)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm') + theme_bw(base_size = 14) + labs(x='SLD (mm)', y=expression(log[2](exoDNA~ng/mL + 1))) + theme(legend.text=element_text(size=9))
plot(SLDvexo)
plot(SLDvexong)

## CFDNA
SLDvcf <- ggplot(met, aes(x=SLD, y=cfDNA)) + geom_point(aes(color=Metastasis, shape=Metastasis), size=2.5) + stat_smooth(method='glm') + theme_bw(base_size=25) + labs(x='SLD (mm)', y=expression(ct*italic(KRAS)~MAF)) + theme(legend.text=element_text(size=15)) + annotate("text", x=60, y=55, label="f(x) = -3.154 + 0.1134x \n p = 0.000848", size=10)
SLDvcfng <- ggplot(met, aes(x=SLD, y=log2cfng)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm')+ theme_bw(base_size=14) + labs(x='SLD (mm)', y=expression(log[2](cfDNA~ng/mL + 1))) + theme(legend.text=element_text(size=9))
plot(SLDvcf)
plot(SLDvcfng)

## vs one another
exovcf <- ggplot(met, aes(x=log2exo, y=log2cf)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm')+ theme_bw(base_size=14) + labs(x=expression(log[2](exoDNA~MAF + 1)), y=expression(log[2](cfDNA~MAF + 1))) + theme(legend.text=element_text(size=9))
plot(exovcf)
```

#### Box/Jitter

```{r plotsJitterBoxmet, echo=FALSE}
comp <- list(c('Liver', 'Lung'), c('Liver', 'Peritoneal'), c('Lung', 'Peritoneal'))
## Box/Jitter
## Met vs 
## Metvexo <- ggplot(met, aes(x=Metastasis, y=log10exo, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=.01) + geom_jitter()+ theme_bw(base_size=25) + guides(color = FALSE, shape=FALSE) + labs(y=expression(log[10](exoDNA~MAF +1))) + stat_compare_means(comparisons = comp) + stat_compare_means(label.y=2.5)
Metvexol <- ggplot(met, aes(x=Metastasis, y=exop1, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=.01) + geom_jitter(size=3)+ theme_bw(base_size=25) + guides(color = FALSE, shape=FALSE) + labs(y=expression(exo*italic(KRAS)~MAF~(log[10]))) + stat_compare_means(comparisons = comp, method='wilcox.test', size=1.5) + stat_compare_means(label.y=2.25, size=8) + scale_y_continuous(trans='log10') + annotation_logticks(sides = 'lr')
## Metvcf <- ggplot(met, aes(x=Metastasis, y=log10cf, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=0.01) + geom_jitter()+ theme_bw(base_size = 25)+ guides(color = FALSE, shape=FALSE) + labs(y=expression(log[10](cfDNA~MAF +1))) + stat_compare_means(comparisons = comp) + stat_compare_means(label.y=2.5)
Metvcfl <- ggplot(met, aes(x=Metastasis, y=cfp1, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=0.01) + geom_jitter(size=3)+ theme_bw(base_size = 25)+ guides(color = FALSE, shape=FALSE) + labs(y=expression(ct*italic(KRAS)~MAF~(log[10]))) + stat_compare_means(comparisons = comp, size=2) + stat_compare_means(label.y=2.25, size=8) + scale_y_continuous(trans='log10')  + annotation_logticks(sides='lr')
MetvSLD <- ggplot(met, aes(x=Metastasis, y=SLD, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=.01) + geom_jitter(size=3)+ theme_bw(base_size=25) + guides(color = FALSE, shape=FALSE) + labs(y='SLD (mm)') +stat_compare_means(comparisons = comp, size=1.5) + stat_compare_means(label.y=260, size=8)
plot(Metvexol)
plot(Metvcfl)
plot(MetvSLD)

## Prog/none
progvexo <- ggplot(met, aes(x=factor(Progression), y=log2exo)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size = 14) + labs(y=expression(log[2](exoDNA~MAF + 1)), x='Progression') + scale_x_discrete(labels=c("0" = "No Progression", "1" = "Progression"))+ theme(legend.text=element_text(size=9))
plot(progvexo)
progvexong <- ggplot(met, aes(x=factor(Progression), y=log2exong)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size = 14) + labs(y=expression(log[2](exoDNA~ng/mL + 1)), x='Progression') + scale_x_discrete(labels=c("0" = "No Progression", "1" = "Progression"))+ theme(legend.text=element_text(size=9))
plot(progvexong)
progvcf <- ggplot(met, aes(x=factor(Progression), y=log2cf)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size = 14) + labs(y=expression(log[2](cfDNA~MAF + 1)), x='Progression') + scale_x_discrete(labels=c("0" = "No Progression", "1" = "Progression"))+ theme(legend.text=element_text(size=9))
plot(progvcf)
progvcfng <- ggplot(met, aes(x=factor(Progression), y=log2cfng)) + geom_boxplot(alpha=.01, width=.5) + geom_jitter(aes(color=Metastasis, shape=Metastasis), width=.25)+ theme_bw(base_size = 14) + labs(y=expression(log[2](exoDNA~ng/mL + 1)), x='Progression') + scale_x_discrete(labels=c("0" = "No Progression", "1" = "Progression")) + theme(legend.text=element_text(size=9))
plot(progvcfng)

## ECOG
comp2 <- list(c('0','2'),  c('1', '2'), c('2','3'))
comptest <- list(c('0','2'),  c('1', '2'), c('2','3'), c('0','1'), c('0', '3'), c('1','3'))
ecogvexol <- ggplot(subset(met, !is.na(ecog)), aes(x=factor(ecog), y=exop1, color=ecog, na.rm=TRUE)) + geom_boxplot(alpha=.01) + geom_jitter(size=3)+ theme_bw(base_size=25) + guides(color = FALSE, shape=FALSE) + labs(y=expression(exo*italic(KRAS)~MAF~(log[10])), x="ECOG") + stat_compare_means(comparisons = comptest, method='wilcox.test', size=1.5) + stat_compare_means(label.y=2.25, size=8) + scale_y_continuous(trans='log10') + annotation_logticks(sides = 'lr') + theme(axis.text = axii.text)
ecogvcfl <- ggplot(subset(met, !is.na(ecog)), aes(x=factor(ecog), y=cfp1, color=ecog, na.rm=TRUE)) + geom_boxplot(alpha=.01) + geom_jitter(size=3)+ theme_bw(base_size=25) + guides(color = FALSE, shape=FALSE) + labs(y=expression(ct*italic(KRAS)~MAF~(log[10])), x="ECOG") + stat_compare_means(label.y=2.25, size=8) + scale_y_continuous(trans='log10') + annotation_logticks(sides = 'lr') + theme(axis.text=axii.text) +stat_compare_means(comparisons = comptest, method='wilcox.test', size=1.5)
plot(ecogvexol)
plot(ecogvcfl)
```

### LA

#### GLM

```{r plotsGLMLA, echo=FALSE}
## GLM
## exoDNA
laSLDvexo <- ggplot(la, aes(x=SLD, y=exoDNA)) + geom_point(size=2.5) + stat_smooth(method='glm') + theme_bw(base_size = 25) + labs(x='SLD (mm)', y=expression(exo*italic(KRAS)~MAF)) + theme(legend.text=element_text(size=15))
laSLDvexong <- ggplot(la, aes(x=SLD, y=log2exong)) + geom_point() + stat_smooth(method='glm') + theme_bw(base_size = 14) + labs(x='SLD (mm)', y=expression(log[2](exoDNA~ng/mL + 1))) + theme(legend.text=element_text(size=9))
plot(laSLDvexo)
plot(laSLDvexong)

## CFDNA
laSLDvcf <- ggplot(la, aes(x=SLD, y=cfDNA)) + geom_point(size=2.5) + stat_smooth(method='glm') + theme_bw(base_size=25) + labs(x='SLD (mm)', y=expression(ct*italic(KRAS)~MAF)) + theme(legend.text=element_text(size=15))
laSLDvcfng <- ggplot(la, aes(x=SLD, y=log2cfng)) + geom_point() + stat_smooth(method='glm')+ theme_bw(base_size=14) + labs(x='SLD (mm)', y=expression(log[2](cfDNA~ng/mL + 1))) + theme(legend.text=element_text(size=9))
plot(laSLDvcf)
plot(laSLDvcfng)

## vs one another
laexovcf <- ggplot(la, aes(x=log2exo, y=log2cf)) + geom_point(aes(color=Metastasis, shape=Metastasis)) + stat_smooth(method='glm')+ theme_bw(base_size=14) + labs(x=expression(log[2](exoDNA~MAF + 1)), y=expression(log[2](cfDNA~MAF + 1))) + theme(legend.text=element_text(size=9))
plot(laexovcf)
```

#### Box/Jitter

```{r plotsBJLA, echo=FALSE}

Metvexol <- ggplot(met, aes(x=Metastasis, y=exop1, color=Metastasis, shape=Metastasis)) + geom_boxplot(alpha=.01) + geom_jitter(size=3)+ theme_bw(base_size=25) + guides(color = FALSE, shape=FALSE) + labs(y=expression(exo*italic(KRAS)~MAF~(log[10]))) + stat_compare_means(comparisons = comp, method='wilcox.test', size=1.5) + stat_compare_means(label.y=2.25, size=8) + scale_y_continuous(trans='log10') + annotation_logticks(sides = 'lr')

```

### Comparing Resectables with Unresectables

#### exoDNA and cfDNA

```{r plotsCohorts, echo=FALSE}
exostage <- ggplot(NULL)
```
